---
title: "fitting2"
author: "Julian Gottfried"
date: "2026-08-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r}
library(tidyverse)
setwd("~/Desktop/CME")
```

## Functions
```{r}
singlerun <- function(alpha, f, R0, G, S, I) {
    Ss <- matrix(0, nrow = 168, ncol = G)
    Is <- matrix(0, nrow = 168, ncol = G)
    t <- rep(0, 168)
    g <- 0
    
    while (g < G) {
        pop = S + I
        if (pop > 0 & pop <= 168) {
            if (g >= t[pop]) {
                t[pop] <- t[pop] + 1
                Ss[pop, t[pop]] <- S
                Is[pop, t[pop]] <- I
            }
        }
        if (S + I == 0) { foi = 0
        } else {foi <- R0 * S * I / (S + I)}
        rate <- alpha + foi + S + I
        g <- g - 1 / rate * log(1 - runif(1))
        pick <- runif(1)
        if (pick < alpha * (1 - f) / rate) { S <- S + 1
        } else if (pick < alpha / rate) { I <- I + 1
        } else if (pick < (alpha + foi) / rate) { 
            S <- S - 1
            I <- I + 1
        } else if (pick < (alpha + foi + S) / rate) { S <- S - 1
        } else {I <- I - 1}
    }
    list(Ss, Is)
}

fitting <- function(statistics, counts) {
    factor <- 0
    loglik <- rep(0, 100*100)
    for (j in unique(statistics$N)) {
        if (j == 0) next
        slimdata <- statistics %>% filter(N == j)
        slimI <- as.numeric(unname(table(factor(slimdata$I, levels = 0:(O-1)))))
        factor <- factor + log(factorial(sum(slimI))) - sum(log(factorial(slimI)))
        slimfreq <- counts %>% filter(N == j)
        loglik <- loglik + apply(slimfreq[, 1:O], 1, \(x) sum(slimI * log(x)))
    }
    
    # https://journals.sagepub.com/doi/pdf/10.1177/1536867X0700700305
    
    hat <- which.max(loglik)
    
    profiledf <- sapply(f, \(x) max(loglik[round(slimfreq$f, 2) == round(x, 2)]))
    finterval <- 2 * (max(loglik) - profiledf) <= qchisq(0.95, 1)

    profiledR0 <- sapply(R0, \(x) max(loglik[round(slimfreq$R0, 2) == round(x, 2)]))
    R0interval <- 2 * (max(loglik) - profiledR0) <= qchisq(0.95, 1)

    returner <- rep(0, 8)
    
    returner[1] <- slimfreq$f[hat]
    returner[2] <- range(f[finterval])[1]
    returner[3] <- range(f[finterval])[2]

    returner[4] <- slimfreq$R0[hat]
    returner[5] <- range(R0[R0interval])[1]
    returner[6] <- range(R0[R0interval])[2]

    returner[7] <- pchisq(2 * (max(loglik) - max(loglik[slimfreq$R0 == 0])), 
                            1, lower.tail = F, log.p = F) / 2
    
    returner[8] <- max(loglik)
    
    returner
}
```

## Hyperparams
```{r}
N <- 42
O <- 50

f <- seq(from = 0, to = 0.495, by = 0.005)
R0 <- seq(from = 0, to = 4.95, by = 0.05)
```

## CME data
```{r}
cmemat <- read.table("eitan_data/host_cme_matrix.txt", header = T, sep = "\t")
cmes <- colnames(cmemat %>% select(-host_id))
cmemat$aid <- str_split(cmemat$host_id, "_", simplify = T)[, 1]
cmemat$genome <- str_split(cmemat$host_id, "_", simplify = T)[, 2]
hostdat <-  read.table("eitan_data/hosts.txt", header = T, sep = "\t")
cmemat <- left_join(cmemat, hostdat %>% select(species, genus, family, host_id), by = "host_id")
cmetab <- read.table("eitan_data/cme_table.txt", header = T, sep = "\t")
cmedat <- cmemat %>% 
    group_by(aid, species) %>% 
    summarize_at(cmes, \(x) sum(x) != 0) %>% 
    select(-species) %>% 
    pivot_longer(cols = -aid, names_to = "cme") %>% 
    group_by(cme, aid) %>% 
    summarize(count = sum(value), .groups = "drop")
```

## N and I data
```{r}
NI <- data.frame(matrix(0, N * length(cmes), 3))
colnames(NI) <- c("cme","N","I")
for (i in 1:length(cmes)) {
    print(i)
    fams <- cmemat %>% filter(.[[cmes[i]]]) %>% pull(family) %>% unique
    Nstats <- suppressMessages(cmemat %>% 
        filter(family %in% fams) %>% 
        group_by(aid) %>% 
        select(aid, species) %>%
        distinct %>% 
        summarize(N = n()))
    Istats <- suppressMessages(cmemat %>% filter(.[[cmes[i]]]) %>% group_by(aid, species) %>% 
        summarize(n = n()) %>% summarize(I = n()))
    statistics <- left_join(Nstats, Istats, by = "aid")
    statistics <- rbind(statistics,
                        data.frame(aid = unique(hostdat$aid), N = 0, I = NA)[!(unique(hostdat$aid) %in% statistics$aid), ])
    statistics[is.na(statistics)] <- 0
    NI[N * (i - 1) + (1:N), ] <- statistics %>% mutate(cme = cmes[i]) %>% select(cme, N, I)
}
```

## Estimate alpha
```{r}
alphas <- NI %>% mutate(cme = factor(NI$cme, levels = unique(NI$cme))) %>% 
    group_by(cme) %>% summarize(alpha = round(mean(N, 0))) %>% 
    mutate(cme = varhandle::unfactor(cme))
alphalist <- sort(unique(alphas$alpha))

alphas %>% arrange(-alpha) %>% mutate(cme = row_number()) %>% 
    ggplot(aes(x = alpha)) +
    geom_histogram(fill = "white", color = "black") +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))
```

## Convert simulations to correct format
```{r}
converter <- function(i) {
    tmpcounts <- data.frame(t(read.csv(paste0("results2/counts_", i,".csv"), header = F)))
    tmpcounts$alpha <- alphalist[i]
    tmpcounts$f <- 0
    tmpcounts$R0 <- 0
    for (j in 1:100) {
        for (k in 1:100) {
            tmpcounts$f[100 * (j - 1) + k] <- f[j]
            tmpcounts$R0[100 * (j - 1) + k] <- R0[k]
        }
    }
    shapely <- matrix(0, 10000 * 168, 54)
    for (j in 1:168) {
        shapely[c(10000 * (j - 1) + (1:10000)), 51] <- j
        shapely[c(10000 * (j - 1) + (1:10000)), c(1:50, 52:54)] <- 
            as.matrix(tmpcounts[, c(50 * (j - 1) + (1:50), 8401:8403)])
    }
    shapely <- data.frame(shapely)
    colnames(shapely)[51:54] <- c("N","alpha","f","R0")
    write_csv(shapely, paste0("results3/counts_", i,".csv"))
}
for (i in 1:43) {
    print(i)
    converter(i)
}
```

## Retrodiction: setup
```{r}
testalpha <- 33
testcounts <- read_csv(paste0("results2/counts_", which(alphalist == testalpha), ".csv"))
testcounts[, 1:O] <- testcounts[, 1:O] + 1 / 10
rowsums <- rowSums(testcounts[, 1:O])
testcounts[, 1:O] <- testcounts[, 1:O] %>% 
    mutate_all(\(x) x / rowsums)

condN <- rpois(N, testalpha)
samplepars <- rbind(testcounts %>% filter(R0 == 0) %>% 
                        select(f, R0) %>% sample_n(100),
                    testcounts %>% filter(R0 > 0) %>% 
                        select(f, R0) %>% sample_n(100))
```

## Retrodiction: fitting
```{r}
testret <- matrix(0, nrow=200, ncol=9)
for (i in 1:200) {
    print(i)
    sim <- apply((testcounts %>% filter(f == samplepars[i, ]$f, R0 == samplepars[i, ]$R0))[condN, 1:O],
                 1, \(x)  which.max(rmultinom(1, 1, x)) - 1)
    
    statistics <- data.frame(N = condN, I = sim)
    
    testret[i, 1:8] <- fitting(statistics, testcounts)
    
    fj <- statistics %>% filter(N > 0) %>% 
        mutate(f = I / N) %>% pull(f) %>% mean
    pi <- 1 - (1 - fj) ^ statistics$N
    if (sum(pi) == 0) { testret[i, 9] <- 1
    } else {
        obs <- sum(statistics$I > 0)
        testret[i, 9] <- poisbinom::ppoisbinom(obs, pi)
    }
}
```

## Retrodiction: viz
```{r}
explore <- data.frame(cbind(testret, samplepars))
colnames(explore)[1:9] <- c("fhat","fl","fu","R0hat","R0l","R0u","pval4","loglik","pval3")

explore %>% 
    filter(!(R0l == 0 & R0u == 9.9)) %>% 
    ggplot(aes(x = R0, y = (R0l + R0u) / 2)) +
    geom_linerange(aes(ymin = R0l, ymax = R0u), alpha = 0.15) +
    geom_point(alpha = 0.9, shape = 1) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))

explore %>% 
    filter(!(R0l == 0 & R0u == 9.9)) %>% 
    ggplot(aes(x = f, y = (fl + fu) / 2)) +
    geom_linerange(aes(ymin = fl, ymax = fu), alpha = 0.15) +
    geom_point(alpha = 0.9, shape = 1) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))

explore %>%
    mutate(is0 = ifelse(R0 == 0, "R0 = 0", "R0 > 0")) %>% 
    ggplot(aes(x = pval4)) +
    geom_histogram(color = "black", fill = "white") + 
    facet_wrap(~is0) +
    xlab("p-value") +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))

# explore %>%
#     mutate(is0 = ifelse(R0 == 0, "R0 = 0", "R0 > 0")) %>% 
#     ggplot(aes(x = pval3)) +
#     geom_histogram(color = "black", fill = "white") + 
#     facet_wrap(~is0) +
#     xlab("p-value") +
#     theme_classic() +
#     theme(text = element_text(size = 12, family = "mono"))
```

## CME fitting
```{r}
fits <- data.frame(cme = cmes, 
                   fhat = 0, fl = 0, fu = 0,
                   R0hat = 0, R0l = 0, R0u = 0,
                   pval = 1,
                   loglik = -Inf)
idx <- 1
for (i in 1:43) {
    tmpalpha <- alphalist[i]
    counts <- suppressMessages(read_csv(paste0("results2/counts_", i, ".csv")))
    counts[, 1:O] <- counts[, 1:O] + 1 / 10
    divider <- rowSums(counts[, 1:O])
    counts[, 1:O] <- apply(counts[, 1:O], 2, \(x) x / divider)
    
    if (i == 43) { tmpcmes <- alphas$cme[alphas$alpha >= tmpalpha]
    } else tmpcmes <- alphas$cme[alphas$alpha == tmpalpha]
    
    for (j in 1:length(tmpcmes)) {
        print(idx)
        tmpNI <- NI %>% filter(cme == tmpcmes[j])
        fits[idx, ] <- c(tmpcmes[j], fitting(tmpNI, counts))
        idx <- idx + 1
    }
}
write_csv(fits, "fits2.csv")
```

## Viz
```{r}
fits <- read_csv("fits2.csv")
fits <- fits %>% 
    mutate(across(-cme, as.numeric))

NI %>% group_by(cme) %>% summarize(test = sum(I)/sum(N)) %>% 
    right_join(fits, by = "cme") %>% 
    mutate(failed = R0u == 4.95 & R0l == 0,
           test = round(test, 3)) %>% 
    group_by(test) %>% summarize(pfail = mean(failed)) %>% 
    ggplot(aes(x = test, y = pfail)) +
    geom_point(size = 2.5, shape = 10) +
    theme_classic() +
    xlab("fraction of observations infected") + ylab("fitting failure") +
    theme(text = element_text(size = 12, family = "mono"))

fits %>% 
    arrange(-(R0l + R0u) / 2) %>% 
    mutate(cme = row_number(), sig = pval >= 0.05) %>% 
    ggplot(aes(x = cme, y = (R0l + R0u) / 2, color = sig)) +
    geom_linerange(aes(ymin = R0l, ymax = R0u), alpha = 0.15) +
    geom_point(alpha = 0.8, shape = 3) +
    scale_color_manual(name = "",
                       labels = c("significant","non-significant"),
                       values = c("#910027", "#818ade")) +
    theme_classic() +
    ylab("R0") +
    theme(text = element_text(size = 12, family = "mono"),
          legend.position = "bottom")

fits %>% 
    arrange(-(fl + fu) / 2) %>% 
    mutate(cme = row_number()) %>% 
    ggplot(aes(x = cme, y = (fl + fu) / 2)) +
    geom_linerange(aes(ymin = fl, ymax = fu), alpha = 0.15) +
    geom_point(alpha = 0.8, shape = 3) +
    theme_classic() +
    ylab("f") +
    theme(text = element_text(size = 12, family = "mono"))

fits %>% 
    arrange(-pval) %>% 
    mutate(cme = row_number()) %>% 
    ggplot(aes(x = cme, y = pval)) +
    geom_point(alpha = 0.8, shape = 3) +
    geom_hline(yintercept = 0.05, color = "red", linetype = 2) +
    theme_classic() +
    ylab("p-value") +
    theme(text = element_text(size = 12, family = "mono"))
```

## Examining specific CMEs
```{r}
tester <- "c4"
parfit <- fits[fits$cme == tester, c("fhat", "R0hat")]
testcounts <- read_csv(paste0("results2/counts_",
                          which(alphalist == min(alphas$alpha[alphas$cme == tester], 92)),
                          ".csv"))
testcounts <- testcounts %>% filter(round(f, 3) == with(parfit, round(fhat, 3)), 
                                    round(R0, 2) == with(parfit, round(R0hat, 2)))
gatherNI <- colSums(testcounts[NI[NI$cme == tester, "N"], ])
gatherNI <- gatherNI / sum(gatherNI)

cmedat %>%
    filter(cme == tester) %>%
    pull(count) %>% 
    factor(levels = 0:(O-1)) %>% 
    table %>% 
    data.frame %>% 
    rename(count = '.', freq = Freq) %>% 
    mutate(count = as.numeric(count) - 1, 
           freq = freq / sum(freq)) %>% 
    ggplot(aes(x = count, y = freq)) +
    geom_col(width = 1, color = "black", fill = "white")+
    ylim(0, 1) + xlim(-1, 20) +
    theme_classic() +
    ggtitle(tester) +
    theme(text = element_text(size = 12, family = "mono"),
          legend.position = "bottom")

data.frame(count = 0:(O - 1), freq = gatherNI[1:O]) %>% 
    ggplot(aes(x = count, y = freq)) +
    geom_col(width = 1, color = "black", fill = "white")+
    ylim(0, 1) + xlim(-1, 20) +
    theme_classic() +
    ggtitle("best fit") +
    theme(text = element_text(size = 12, family = "mono"),
          legend.position = "bottom")
```

## Permutation --> Model: step 1
```{r}
pval1 <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    fj <- suppressMessages(cmemat %>% mutate(inq = cmemat[[cmes[i]]]) %>% 
        group_by(species) %>% summarize(f = mean(inq)))
    pi <- left_join(hostdat, fj, by = "species") %>%
        group_by(aid) %>%
        select(aid, species, f) %>% 
        distinct %>% 
        summarize(pi = 1 - prod(1 - f)) %>%
        pull(pi)
    obs <- cmemat %>% mutate(inq = cmemat[[cmes[i]]]) %>% 
        group_by(aid) %>% summarize(present = sum(inq) > 0) %>% 
        pull(present) %>% sum
    pval1[i] <- poisbinom::ppoisbinom(obs, pi)
}
data.frame(cme = 1:length(cmes), pval = sort(pval1, decreasing = T)) %>% 
    ggplot(aes(x = cme, y = pval)) +
    geom_point(alpha = 0.8, shape = 3) +
    geom_hline(yintercept = 0.05, color = "red", linetype = 2) +
    theme_classic() +
    ylab("p-value") +
    theme(text = element_text(size = 12, family = "mono"))
```

## Permutation --> Model: step 2
```{r}
pval2 <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    fj <- suppressMessages(cmemat %>% mutate(inq = cmemat[[cmes[i]]]) %>% 
        group_by(species, aid) %>% summarize(inq = sum(inq) > 0) %>% 
        summarize(f = mean(inq)))
    pi <- left_join(hostdat, fj, by = "species") %>%
        group_by(aid) %>%
        select(aid, species, f) %>% 
        distinct %>% 
        summarize(pi = 1 - prod(1 - f)) %>%
        pull(pi)
    obs <- cmemat %>% mutate(inq = cmemat[[cmes[i]]]) %>% 
        group_by(aid) %>% summarize(present = sum(inq) > 0) %>% 
        pull(present) %>% sum
    pval2[i] <- poisbinom::ppoisbinom(obs, pi)
}
data.frame(cme = 1:length(cmes), pval = sort(pval2, decreasing = T)) %>% 
    ggplot(aes(x = cme, y = pval)) +
    geom_point(alpha = 0.8, shape = 3) +
    geom_hline(yintercept = 0.05, color = "red", linetype = 2) +
    theme_classic() +
    ylab("p-value") +
    theme(text = element_text(size = 12, family = "mono"))
```

## Permutation --> Model: step 3
```{r}
pval3 <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    fj <- NI[NI$cme == cmes[i], ] %>% filter(N > 0) %>% 
        mutate(f = I / N) %>% pull(f) %>% mean
    pi <- 1 - (1 - fj) ^ NI[NI$cme == cmes[i], "N"]
    obs <- cmemat %>% mutate(inq = cmemat[[cmes[i]]]) %>% 
        group_by(aid) %>% summarize(present = sum(inq) > 0) %>% 
        pull(present) %>% sum
    pval3[i] <- poisbinom::ppoisbinom(obs, pi)
}
data.frame(pval = pval3, alpha = alphas$alpha) %>% 
    arrange(-pval) %>% 
    mutate(cme = row_number()) %>% 
    ggplot(aes(x = cme, y = pval)) +
    geom_point(alpha = 0.8, shape = 3) +
    geom_hline(yintercept = 0.05, color = "red", linetype = 2) +
    theme_classic() +
    ylab("p-value") +
    theme(text = element_text(size = 12, family = "mono"))
```

## All steps
```{r}
pval0 <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    pval0[i] <- cmetab[cmetab$cseg == cmes[i], ]$p_subjects
}
pval4 <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    pval4[i] <- fits[fits$cme == cmes[i], ]$pval
}
pval_labs <- c("pval0" = "Permutation test",
               "pval1" = "Parametric version",
               "pval2" = "Aggregated at species level",
               "pval3" = "Species made exchangeable",
               "pval4" = "Model-based LRT")
data.frame(cme = cmes, 
           pval0 = pval0, pval1 = pval1, 
           pval2 = pval2, pval3 = pval3,
           pval4 = pval4,
           alpha = alphas$alpha) %>% 
    left_join(fits %>% select(cme, fu), by = "cme") %>% 
    arrange(-pval4) %>% mutate(cme = row_number()) %>% 
    pivot_longer(pval0:pval4) %>% 
    ggplot(aes(x = cme, y = value, color = fu == 0)) +
    geom_point(alpha = 0.8, shape = 3) +
    scale_color_manual(values = c("black", "grey"), 
                       labels = c("f = 0","f > 0"),
                       name = "") +
    geom_hline(yintercept = 0.05, color = "red", linetype = 2) +
    theme_classic() +
    ylab("p-value") +
    theme(text = element_text(size = 12, family = "mono")) +
    facet_wrap(~name, ncol = 1, labeller = labeller(name = pval_labs)) +
    guides(color = guide_legend(override.aes = list(shape = 15)))
```

## Failure of Poisson(alpha) for data
```{r}
examineerror <- NI %>% group_by(cme) %>% 
    summarize(Ealpha = mean(N), Valpha = var(N)) %>% 
    mutate(ratio = Ealpha/Valpha)
examineerror %>%
    ggplot(aes(x = Ealpha, y = Valpha)) +
    geom_point(alpha = 0.5) +
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = 2) +
    theme_classic() +
    xlab("E[alpha]") + ylab("Var[alpha]") +
    theme(text = element_text(size = 12, family = "mono"))
examineerror %>% filter(ratio > 0.75) %>% select(cme, Ealpha)
```

## Significance retrodiction under fitted parameters
```{r}
sigs <- data.frame(cme = cmes, pval3 = 1, pval4 = 0.5)
idx <- 1
for (i in 1:43) {
    tmpalpha <- alphalist[i]
    counts <- suppressMessages(read_csv(paste0("results2/counts_", i, ".csv")))
    counts[, 1:O] <- counts[, 1:O] + 1 / 10
    divider <- rowSums(counts[, 1:O])
    counts[, 1:O] <- apply(counts[, 1:O], 2, \(x) x / divider)
    
    if (i == 43) { tmpcmes <- alphas$cme[alphas$alpha >= tmpalpha]
    } else tmpcmes <- alphas$cme[alphas$alpha == tmpalpha]
    
    for (j in 1:length(tmpcmes)) {
        print(idx)
        tmpN <- NI %>% filter(cme == tmpcmes[j]) %>% filter(N > 0) %>% pull(N)

        sigs[idx, 1] <- as.character(tmpcmes[j])
        
        tmppar <- fits %>% filter(cme == tmpcmes[j]) %>% 
            mutate(f = round((fl + fu) / 2, 2),
                   R0 = round((R0l + R0u) / 2, 1)) %>% 
            select(f, R0)
        
        sim <- apply((counts %>% filter(round(f, 2) == tmppar$f, round(R0, 1) == tmppar$R0))[tmpN, 1:O],
                 1, \(x)  which.max(rmultinom(1, 1, x)) - 1)
    
        statistics <- data.frame(N = tmpN, I = sim)
        
        sigs[idx, 3] <- fitting(statistics, counts)[7]
        
        fj <- statistics %>% mutate(f = I / N) %>% pull(f) %>% mean
        pi <- 1 - (1 - fj) ^ statistics$N
        if (sum(pi) > 0) {
            obs <- sum(statistics$I > 0)
            sigs[idx, 2] <- poisbinom::ppoisbinom(obs, pi)
        }
        idx <- idx + 1
    }
}
examine <- fits %>% mutate(f = round((fl + fu) / 2, 2),
                R0 = round((R0l + R0u) / 2, 1)) %>% 
    select(cme, f, R0, pval4 = pval) %>% 
    left_join(data.frame(cme = cmes, pval3 = pval3), by = "cme") %>% 
    left_join(sigs %>% rename(retro3 = pval3, retro4 = pval4), by = "cme")
examine %>%
    mutate(is0 = ifelse(R0 == 0, "R0 = 0", "R0 > 0")) %>% 
    filter(f > 0) %>% 
    ggplot(aes(x = retro4)) +
    geom_histogram(color = "black", fill = "white") + 
    facet_wrap(~is0) +
    xlab("p-value") +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))
```

## Sim I vs. real I
```{r}
Isim <- list()
idx <- 1
for (i in 1:43) {
    tmpalpha <- alphalist[i]
    counts <- suppressMessages(read_csv(paste0("results2/counts_", i, ".csv")))
    counts[, 1:O] <- counts[, 1:O] + 1 / 10
    divider <- rowSums(counts[, 1:O])
    counts[, 1:O] <- apply(counts[, 1:O], 2, \(x) x / divider)
    
    if (i == 43) { tmpcmes <- alphas$cme[alphas$alpha >= tmpalpha]
    } else tmpcmes <- alphas$cme[alphas$alpha == tmpalpha]
    
    for (j in 1:length(tmpcmes)) {
        print(idx)
        tmpN <- NI %>% filter(cme == tmpcmes[j]) %>% filter(N > 0) %>% pull(N)

        # sigs[idx, 1] <- as.character(tmpcmes[j])
        
        tmppar <- fits %>% filter(cme == tmpcmes[j]) %>% 
            mutate(f = round((fl + fu) / 2, 2),
                   R0 = round((R0l + R0u) / 2, 1)) %>% 
            select(f, R0)
        
        sim <- apply((counts %>% filter(round(f, 2) == tmppar$f, round(R0, 1) == tmppar$R0))[tmpN, 1:O],
                 1, \(x)  apply(rmultinom(100, 1, x), 2, which.max) - 1)

        Isim[[as.character(tmpcmes[j])]] <- sim
        
        idx <- idx + 1
    }
}
summar <- map2(Isim, names(Isim), \(x, y) {
    tmpN <- NI %>% filter(cme == y) %>% pull(N)
    tmpN <- tmpN[tmpN > 0]
    tmpstat <- apply(x, 1, \(row) mean(row / tmpN))
    c(mean(tmpstat), quantile(tmpstat, c(0.055, 0.5, 0.945)))}) %>% 
    bind_rows %>% t %>% data.frame
colnames(summar) <- c("mean","q1","q2","q3")
summar$cme <- rownames(summar)
summar <- NI %>% group_by(cme) %>% 
    summarize(f = mean(I / N, na.rm = T)) %>% 
    left_join(summar, by = "cme")
summar %>% ggplot(aes(x = f, y = q2)) +
    geom_point(alpha = 0.25) +
    geom_linerange(aes(ymin = q1, ymax = q3), alpha = 0.25) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic()
problems <- summar %>% filter(q1 == 0, q2 == 0, q3 == 0) %>% pull(cme)
mean((fits %>% filter(fu == 0) %>% pull(cme)) %in% problems)
```

## Plot for proposal
```{r}
fits %>% 
    filter(!(R0u == 4.95 & R0l == 0)) %>% 
    mutate(R0 = (R0l + R0u) / 2) %>% 
    ggplot(aes(x = R0)) +
    geom_bar(fill = "black") +
    theme_classic() +
    ylab("count") + xlab(expression(R["0"])) +
    theme(text = element_text(size = 12, family = "mono"))
fits %>% 
    filter(!(R0u == 4.95 & R0l == 0)) %>% 
    mutate(R0 = (R0l + R0u) / 2) %>% 
    ggplot(aes(x = R0, fill = pval > 0.05)) +
    geom_bar(position = position_stack()) +
    scale_fill_manual(name = "",
                       labels = c("p < 0.05","n.s."),
                       values = c("#818ade","black")) +
    theme_classic() +
    ylab("count") + xlab(expression(R["0"])) +
    theme(text = element_text(size = 12, family = "mono"))
fits %>% 
    arrange(-(R0l + R0u) / 2) %>% 
    filter(!(R0u == 4.95 & R0l == 0)) %>% 
    mutate(cme = row_number()) %>% 
    ggplot(aes(x = cme, y = (R0l + R0u) / 2)) +
    geom_point(shape = 3) +
    geom_linerange(aes(ymin = R0l, ymax = R0u), alpha = 0.4) +
    theme_classic() +
    ylab("R0-within") + xlab("mobile element, rank-ordered") +
    theme(text = element_text(size = 12, family = "mono"))
fits %>% 
    arrange(-(R0l + R0u) / 2) %>% 
    filter(!(R0u == 4.95 & R0l == 0)) %>% 
    mutate(cme = row_number(), sig = pval < 0.05) %>% 
    ggplot(aes(x = cme, y = (R0l + R0u) / 2, color = !sig)) +
    geom_point(shape = 3) +
    geom_linerange(aes(ymin = R0l, ymax = R0u), alpha = 0.6) +
    scale_color_manual(name = "Significance",
                       labels = c("< 0.05","n.s."),
                       values = c("#6972c9", "black")) +
    theme_classic() +
    ylab("R0-within") + xlab("mobile element, rank-ordered") +
    theme(text = element_text(size = 12, family = "mono"),
          legend.position = "bottom")
```

## Fitting success
```{r}
fits %>%
    filter(!(R0u == 4.95 & R0l == 0)) %>%
    mutate(R0 = (R0l + R0u) / 2) %>%
    arrange(-loglik) %>%
    mutate(idx = row_number()) %>% 
    ggplot(aes(x = loglik)) +
    geom_histogram(color = "black", fill = "white", bins = 30) +
    theme_classic() +
    theme(panel.background = element_rect(fill = '#f2f2f2', colour = 'black'),
          text = element_text(size = 12, family = "mono"))
fits %>%
    filter(!(R0u == 4.95 & R0l == 0)) %>%
    mutate(R0 = (R0l + R0u) / 2) %>%
    arrange(-loglik) %>%
    mutate(idx = row_number()) %>% 
    ggplot(aes(x = R0, y = loglik, color = cme == "c4", alpha = pval < 0.05)) +
    geom_point(shape = 1, stroke = 0.75) +
    scale_color_manual(values = c("black","red")) +
    scale_alpha_manual(values = c(0.1, 0.7)) +
    theme_classic() +
    guides(color = "none", alpha = "none") +
    ylab("log lik") + xlab(expression(R["0"])) +
    theme(text = element_text(size = 12, family = "mono"))
```

## Species density distribution
```{r}
hostdat %>% group_by(aid, species) %>% summarize(n=n()) %>% 
    mutate(p = n / n()) %>% ungroup %>% arrange(-p) %>% 
    mutate(p = n() * p / sum(p)) %>% mutate(sp = row_number()) %>% 
    mutate(sim = sort(rlnorm(n(), log(1 / sqrt(1 + var(p))), sqrt(log(1 + var(p)))), decreasing = T)) %>% 
    rename(smooth = sim, data = p) %>% 
    pivot_longer(cols = c(smooth, data)) %>% 
    ggplot(aes(x = sp, y = value, color = name)) +
    geom_point(alpha = 0.05, size = 1) +
    scale_color_manual(values = c("black", "#db48c3"), name = "") +
    theme_classic() +
    ylab("scaled abudance") + xlab("species, rank-ordered") +
    theme(text = element_text(size = 12, family = "mono")) +
    guides(color = guide_legend(override.aes = list(alpha = 1)))

hostdat %>% group_by(aid, species) %>% summarize(n=n()) %>% 
    mutate(p = n / n()) %>% ungroup %>% arrange(-p) %>% 
    mutate(p = n() * p / sum(p)) %>% mutate(sp = row_number()) %>% 
    mutate(sim = sort(rlnorm(n(), log(1 / sqrt(1 + var(p))), sqrt(log(1 + var(p)))), decreasing = T)) %>%
    rename(smooth = sim, data = p) %>% 
    pivot_longer(cols = c(smooth, data)) %>% 
    ggplot(aes(y = name, x = value)) +
    ggridges::geom_density_ridges(stat = "binline", bins = 40, alpha = 0.8, scale = 3,
                                  fill = "black", color = "white") +
    theme_classic() +
    xlab("scaled abundance") + ylab("") +
    xlim(0, 4) +
    theme(text = element_text(size = 12, family = "mono"))

testvar <- hostdat %>% group_by(aid, species) %>% summarize(n=n()) %>% 
    mutate(p = n / n()) %>% ungroup %>% arrange(-p) %>% 
    mutate(p = n() * p / sum(p)) %>% pull(p) %>% var

mu <- log(1 / sqrt(1 + testvar))
sigmasq <- sqrt(log(1 + testvar))
```
