---
title: "fitting"
author: "Julian Gottfried"
date: "2026-07-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set-up
```{r}
library(tidyverse)
setwd("~/Desktop/CME")
```

## Local sim functions
```{r}
singlerun <- function(alpha, f, R0, G, S, I) {
    Ss <- rep(0, G)
    Is <- rep(0, G)
    t <- 0
    g <- 0
    while (g < G) {
        if (g >= t) {
            Ss[t + 1] <- S
            Is[t + 1] <- I
            t <- t + 1
        }
        if (S + I == 0) { foi = 0
        } else {foi <- R0 * S * I / (S + I)}
        rate <- alpha + foi + S + I
        g <- g - 1 / rate * log(1 - runif(1))
        pick <- runif(1)
        if (pick < (1 - f) * alpha / rate) { S <- S + 1
        } else if (pick < alpha / rate) { I <- I + 1
        } else if (pick < (alpha + foi) / rate) { 
            S <- S - 1
            I <- I + 1
        } else if (pick < (alpha + foi + S) / rate) { S <- S - 1
        } else {I <- I - 1}
    }
    c(Ss, Is)
}

plotsingle <- function(result, G) {
    data.frame(time = c(1:G, 1:G), size = result, state = c(rep("S", G), rep("I", G))) %>%
        mutate(state = factor(state, levels = c("S", "I"))) %>%
        ggplot(aes(x = time, y = size)) +
        geom_line() +
        facet_wrap(~state) +
        theme_classic() +
        theme(text=element_text(size=12,family="mono"),
              legend.position = "bottom")
}
```

## Grid parameters
```{r}
alpha <- seq(from=0,by=1,to=99)
f <- seq(from=0,by=0.005,to=0.495)
R0 <- seq(from=0,by=0.1,to=9.9)

N <- 42
O <- 50
```

## Aggregate cluster sims
```{r}
counts <- NULL
for (i in 1:100) {
    print(i)
    countstmp <- data.frame(t(read.table(paste0("JLresults/counts_",i,".csv"), sep=",")))
    countstmp$alpha <- alpha[i]
    countstmp$f <- 0
    countstmp$R0 <- 0
    for (j in 1:100) {
        for (k in 1:100) {
            idx <- 100 * (j - 1) + k
            countstmp$f[idx] <- f[j]
            countstmp$R0[idx] <- R0[k]
        }
    }
    counts <- rbind(counts, countstmp)
}
write_csv(counts, "results.csv")
```

## Convert counts
```{r}
counts <- read_csv("results.csv")
freqs <- counts
pseudo <- 1
freqs[, 1:O] <- freqs[, 1:O] + pseudo
rowsums <- rowSums(freqs[, 1:O])
freqs[, 1:O] <- freqs[, 1:O] %>% 
    mutate_all(\(x) x / rowsums)

write_csv(freqs, "pars.csv", col_names = T)
freqs <- read_csv("pars.csv", col_names = T)

freqs[14000, ] %>% 
    pivot_longer(starts_with("X"), names_to = "bin", values_to = "proportion") %>% 
    mutate(bin = as.numeric(str_remove(bin, "X")) - 1) %>%
    ggplot(aes(x = bin, y = proportion))+
    geom_point()+
    theme_classic()
```

## Get CME data
```{r}
cmemat <- read.table("eitan_data/host_cme_matrix.txt", header = T, sep = "\t")
cmes <- colnames(cmemat %>% select(-host_id))
cmemat$subject <- str_split(cmemat$host_id, "_", simplify = T)[, 1]
cmemat$genome <- str_split(cmemat$host_id, "_", simplify = T)[, 2]
hostdat <-  read.table("eitan_data/hosts.txt", header = T, sep = "\t")
cmemat <- left_join(cmemat, hostdat %>% select(species, genus, family, host_id), by = "host_id")

cmedat <- cmemat %>% 
    group_by(subject, genome) %>% 
    summarize_at(cmes, \(x) sum(x) != 0) %>% 
    select(-genome) %>% 
    pivot_longer(cols = -subject, names_to = "cme") %>% 
    group_by(cme, subject) %>% 
    summarize(count = sum(value), .groups = "drop")
```

## Pre-estimate f
```{r}
# Method that doesn't work
level = "genus"
estf <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    infecting <- cmemat[, level][cmemat[, cmes[i]]]
    estf[i] <- length(unique(infecting)) / length(infecting)
}
ggplot(data.frame(estf),
       aes(x = estf))+
    geom_histogram(color = "white", fill = "black") +
    theme_classic()

# Method that may work
coarse = "family"
fine = "species"
estf <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    infecting <- cmemat[, cmes[i]][cmemat[, fine] %in% 
                hostdat[, fine][hostdat[, coarse] %in% 
                                    unique(cmemat[, coarse][cmemat[, cmes[i]]])]]
    estf[i] <- sum(infecting)/length(infecting)
}
ggplot(data.frame(estf),
       aes(x = estf))+
    geom_histogram(color = "white", fill = "black") +
    xlab("f")+
    theme_classic()

# Snap to grid
closestf <- round(f[estf%/%0.005 + 1], 3)
ggplot(data.frame(closestf),
       aes(closestf))+
    geom_histogram(color = "white", fill = "black") +
    xlab("f")+
    theme_classic()+
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
```

## Do fitting
```{r}
pars <- freqs %>% select(alpha:R0)
fits <- data.frame(cme = cmes, idx = 0, loglik = -Inf)
for (i in 1:length(cmes)) {
    print(i)
    tmpdat <- cmedat %>% filter(cme == cmes[i]) %>% pull(count)
    tmpdat <- as.numeric(unname(table(factor(tmpdat, levels = 0:(O-1)))))
    factor <- log(factorial(N) / prod(factorial(tmpdat)))
    loglik <- rowSums(sweep(log(freqs[, 1:O]), 2, tmpdat, FUN = "*"))
    fits[i, 2:3] <- c(which.max(loglik), factor + max(loglik))
}
write_csv(fits, "fits_genome.csv")
fits <- read_csv("fits_genome.csv")
```

## Do constrained fitting
```{r}
fits0 <- data.frame(cme = cmes, idx = 0, loglik = -Inf)
for (i in 1:length(cmes)) {
    print(i)
    tmpdat <- cmedat %>% filter(cme == cmes[i]) %>% pull(count)
    tmpdat <- as.numeric(unname(table(factor(tmpdat, levels = 0:(O-1)))))
    factor <- log(factorial(N) / prod(factorial(tmpdat)))
    loglik <- apply(freqs[freqs$R0 == 0, 1:O], 1, \(x) sum(tmpdat * log(x)))
    fits0[i, ] <- c(cmes[i], which.max(loglik), factor + max(loglik))
}
```

## Plot parameter spread
```{r}
explore <- pars[fits$idx, ] %>% mutate(cme = cmes)

explore %>% 
    group_by(alpha, f, R0) %>% 
    summarize(count = n()) %>% 
    ungroup %>% 
    mutate(alphaI = alpha * f, alphaS = alpha * (1 - f)) %>%
    pivot_longer(c(alphaS,alphaI,R0), names_to = "parameter") %>% 
    mutate(parameter = factor(parameter, 
                              levels = c("alphaS","alphaI","R0"))) %>% 
    ggplot(aes(x = value)) +
    geom_histogram(color = "white", fill = "black")+
    ggtitle("parameter fits across CMEs") +
    theme_classic() +
    facet_wrap(~parameter, scale="free")+
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

explore %>% 
    group_by(alpha, f, R0) %>%
    summarize(count = n()) %>%
    ungroup %>% 
    mutate(alphaI = alpha * f, alphaS = alpha * (1 - f)) %>% 
    ggplot(aes(x = alphaS, y = R0)) +
    geom_point(aes(size = count), stroke = 0.5, alpha = 1, shape = 1) +
    scale_size(range = c(0.5, 10)) +
    ggtitle("parameter fits across CMEs") +
    theme_classic() +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

sum(explore$R0 != 0)
sum(explore$R0 == 0)
```

## Plot p fit
```{r}
ordering <- order(explore$f, decreasing = T)
data.frame(x = 1:length(cmes), 
           priorf = closestf[ordering],
           estf = explore$f[ordering]) %>% 
    pivot_longer(priorf:estf) %>% 
    ggplot(aes(x = x, y = value, color = name)) +
    scale_color_discrete(palette = c("#E0AB60", "#7DA2C9"), 
                         labels = c("estimated","calculated"),
                         name = "") +
    geom_point() +
    ylab("f") +
    xlab("CME") +
    theme_classic() +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

```

## Plot particular CMEs
```{r}
fits$cme[with(explore, R0 == max(R0))]
fits$cme[with(explore, R0 == 0)]
fits$cme[with(explore, alpha == max(alpha))]
fits$cme[with(explore, alpha == min(alpha))]

tester <- "c86"
cmedat %>%
    filter(cme == tester) %>%
    pull(count) %>% 
    factor(levels = 0:(O-1)) %>% 
    table %>% 
    data.frame %>% 
    rename(count = '.', freq = Freq) %>% 
    mutate(count = as.numeric(count) - 1, 
           freq = freq / sum(freq)) %>% 
    ggplot(aes(x=count, y=freq)) +
    geom_col(width = 1, color = "black", fill = "white")+
    ylim(0,1)+xlim(-1,51)+
    theme_classic() +
    ggtitle(tester) +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
counts[fits$idx[fits$cme == tester], 1:O] %>% 
    pivot_longer(everything(), 
                 names_to = "count", 
                 values_to="freq") %>% 
    mutate(count=as.numeric(str_remove(count,"X"))-1, 
           freq = freq / sum(freq)) %>% 
    ggplot(aes(x=count, y=freq)) +
    geom_col(width = 1, color = "black", fill = "white")+
    ylim(0,1)+xlim(-1,51)+
    theme_classic() +
    ggtitle("best fit") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

testpars <- explore[explore$cme == tester, ]
plotsingle(singlerun(testpars$alpha,
                     testpars$f, 
                     testpars$R0,
                     1000,
                     25, 0), 1000)
```

## Poisson goodness of fit
```{r}
goodness <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    tmpdat <- cmedat %>% filter(cme == cmes[i]) %>% pull(count)
    lambda <- mean(tmpdat)
    tmpdat <- as.numeric(unname(table(factor(tmpdat, levels = 0:(O-1)))))
    expected <- N*dpois(0:(O-1),lambda)
    stat <- sum((tmpdat-expected)^2 / expected)
    df <- O - 2
    goodness[i] <- pchisq(stat, df, lower.tail = F, log.p = T)
}

cmes[exp(goodness)<0.01]

data.frame(cme = cmes, pval = exp(goodness)) %>% 
    arrange(-pval) %>%
    mutate(idx = row_number()) %>% 
    ggplot(aes(x=idx,y=pval))+
    geom_point(alpha = 1, shape = 3)+
    geom_hline(yintercept = 0.01, color="red", linetype=2)+
    theme_classic()+
    xlab("cme")+ylab("p-value")+
    ggtitle("within-subject transmission") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
```

## Negative binomial error
```{r}
nsamp <- 10000
sampling <- freqs[sample(1:nrow(freqs), nsamp), ]
mse <- rep(0, nsamp)
for (i in 1:nsamp) {
    print(i)
    tmpdat <- as.numeric(sampling[i, 1:50])
    asint <- round(1000 * tmpdat, 0)
    p = mean(asint) / var(asint)
    r = mean(asint) * p / (1 - p)
    probs <- dnbinom(0:(O-1),r,p) / sum(dnbinom(0:(O-1),r,p))
    # dens[i] <- gtools::ddirichlet(tmpdat, probs)
    mse[i] <- mean((probs - tmpdat)^2)
}

data.frame(mse = mse) %>% 
    arrange(-mse) %>%
    mutate(sim = row_number()) %>% 
    ggplot(aes(x=sim,y=mse))+
    geom_line()+
    theme_classic()+
    ylim(0,1)+
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
```

## Negative binomial correspondence
```{r}
visit <- sample(1:nrow(pars), 10000)
abut <- cbind(pars[visit, ], mu = 0, sigmasq = 0, r = 0, p = 0)
for (i in 1:nrow(abut)) {
    if (i %% 100 == 1) print(i)
    tmpfreq <- as.numeric(freqs[visit, 1:50][i, ])
    mu <- sum(tmpfreq * (0:49))
    sigmasq <- sum(tmpfreq * (0:49) ^ 2) - mu ^ 2
    p = min(max(mu / sigmasq, 0), 1)
    r = max(floor(mu * p / (1 - p)), 0)
    abut[i, 4:7] <- c(mu, sigmasq, r, p)
}

abut %>% 
    filter(r < 50) %>% 
    ggplot(aes(x = alpha, y = f, fill = p)) +
    geom_tile() +
    theme_classic()
```

## Negative binomial likelihood ratio test
```{r}
lrt <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    tmpdat <- cmedat %>% filter(cme == cmes[i]) %>% pull(count)
    lambda <- mean(tmpdat)
    p = mean(tmpdat) / var(tmpdat)
    r = mean(tmpdat) * p / (1 - p)
    lr <- 2*abs(sum(dnbinom(tmpdat, r, p, log = T)) -
            sum(dpois(tmpdat, lambda, log = T)))
    df <- 1
    lrt[i] <- pchisq(lr, df, lower.tail = F, log.p = T) - log(2)
}

data.frame(cme = cmes, pval = exp(lrt)) %>% 
    arrange(-pval) %>%
    mutate(idx = row_number()) %>% 
    ggplot(aes(x=idx,y=pval))+
    geom_point(alpha = 1, shape = 3)+
    geom_hline(yintercept = 0.01, color="red", linetype=2)+
    theme_classic()+
    xlab("cme")+ylab("p-value")+
    ggtitle("within-subject transmission via LRT") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

data.frame(cme = cmes, good = exp(goodness), lrt = exp(lrt)) %>% 
    arrange(-good) %>%
    mutate(idx = row_number()) %>% 
    pivot_longer(good:lrt, names_to = "test", values_to = "pval") %>% 
    ggplot(aes(x=idx,y=pval, color=test))+
    geom_point(alpha = 0.5, shape = 3)+
    scale_color_discrete(palette=c("#c4933f","#2aa7f5"),
                         labels=c("goodness of fit", "LRT"),
                         name = "test")+
    geom_hline(yintercept = 0.01, color="red", linetype=2)+
    theme_classic()+
    xlab("cme")+ylab("p-value")+
    ggtitle("within-subject transmission") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
```

## Empirical likelihood ratio test
```{r}
emplrt <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    tmpdat <- cmedat %>% filter(cme == cmes[i]) %>% pull(count)
    lr <- 2*abs(as.numeric(fits[fits$cme == cmes[i], 3]) -
                    as.numeric(fits0[fits0$cme == cmes[i], 3]))
    df <- 2
    emplrt[i] <- pchisq(lr, df, lower.tail = F, log.p = T) - log(2)
}

data.frame(cme = cmes, pval = exp(emplrt)) %>% 
    arrange(-pval) %>%
    mutate(idx = row_number()) %>% 
    ggplot(aes(x=idx,y=pval))+
    geom_point(alpha = 1, shape = 3)+
    geom_hline(yintercept = 0.01, color="red", linetype=2)+
    theme_classic()+
    xlab("cme")+ylab("p-value")+
    ggtitle("within-subject transmission via empirical LRT") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

cmes[exp(emplrt) <= 0.01]

data.frame(cme = cmes, good = exp(goodness), lrt = exp(emplrt)) %>% 
    arrange(-good) %>%
    mutate(idx = row_number()) %>% 
    pivot_longer(good:lrt, names_to = "test", values_to = "pval") %>% 
    ggplot(aes(x=idx,y=pval, color=test))+
    geom_point(alpha = 0.5, shape = 3)+
    scale_color_discrete(palette=c("#c4933f","#2aa7f5"),
                         labels=c("goodness of fit", "empirical LRT"),
                         name = "test")+
    geom_hline(yintercept = 0.01, color="red", linetype=2)+
    theme_classic()+
    xlab("cme")+ylab("p-value")+
    ggtitle("within-subject transmission") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
```

## alpha correlations
```{r}
taxa = c("species","genus","family")
prev <- matrix(0, length(cmes), 3, dimnames = list(NULL, taxa))
for (i in 1:length(cmes)) {
    infectable <- cmemat[, taxa][cmemat[, cmes[i]],]
    prev[i, ] <- unlist(lapply(taxa, \(taxon)
                               sum(table(hostdat[,taxon])[
                                   unique(infectable[,taxon])]
                                   ) / nrow(hostdat)))
}

explore %>% bind_cols(prev) %>% 
    pivot_longer(species:family, 
                 names_to="taxon", 
                 values_to="prevalence") %>% 
    ggplot(aes(x = prevalence, y = alpha))+
    geom_point() +
    theme_classic() +
    facet_wrap(~taxon, scales = "free_x")+
    xlab("prevalence of infectable taxa")+ylab("alpha")+
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

# taxon = c("family")
# weighting <- matrix(0, 
#                length(cmes), 
#                length(unique(hostdat[, taxon])), 
#                dimnames = list(NULL, unique(hostdat[, taxon])))
# for (i in 1:length(cmes)) {
#     infectable <- unique(cmemat[, taxon][cmemat[, cmes[i]]])
#     tmpfam <- table(hostdat[, taxon])[infectable]
#     tmpfam <- tmpfam / sum(tmpfam)
#     weighting[i, names(tmpfam)] <- tmpfam * explore[i,]$R0
# }
# colSums(weighting)
```

## Matrix!
```{r}
ndim <- (O + 1) ^ 2

states <- matrix(0, ndim, 2)
k = 1
for (i in 0:O) {
    for (j in 0:O) {
        states[k, ] = c(i, j)
        k = k + 1
    }
}

Q <- matrix(0, ndim, ndim)

alpha1 = 25
alpha2 = 0.5
matR0 = 1

for (i in 1:ndim) {
    print(i)
    for (j in 1:ndim) {
        S = states[i, 1]
        I = states[i, 2]
        dS = states[j, 1] - S
        dI = states[j, 2] - I
        if (dS == 1 & dI == 0) {
            Q[i, j] = alpha1
        } else if (dS == 0 & dI == 1) {
            Q[i, j] = alpha2
        } else if (dS == -1 & dI == 0) {
            Q[i, j] = S
        } else if (dS == 0 & dI == -1) {
            Q[i, j] = I
        } else if (dS == -1 & dI == 1) {
            Q[i, j] = matR0 * S * I / (S + I)
        }
    }
}
for (i in 1:ndim) {
    Q[i, i] = -sum(Q[i, ])
}

pi <- pracma::nullspace(t(Q))
pi <- round(pi / sum(pi), 5)[, 1]

distribution <- data.frame(states) %>% 
    rename(S = X1, I = X2) %>% 
    mutate(pi = pi) %>% 
    group_by(I) %>% 
    summarize(pi = sum(pi))

ggplot(distribution, aes(x = I, y = pi)) +
    geom_point() + 
    theme_classic() +
    ylim(0, 1) +
    theme(text = element_text(size = 12, family = "mono"),
          legend.position = "bottom")
```

## Retrodiction
```{r}
selected <- read_csv("retrodiction/selected.csv", col_names = F)[1:100, ]
colnames(selected) <- c("idx", "alpha", "f", "R0")
selected$idx <- as.integer(selected$idx)

sims <- data.frame()
for (i in 1:100) {
    tmpsim <- data.frame(t(read_csv(paste0("retrodiction/dat_", i, ".csv"), col_names = F)))
    tmpsim$par <- i
    sims <- rbind(sims, tmpsim)
}
sims$rep <- rep(1:100, times = 100)
write_csv(sims, "retrodiction/sims.csv")
sims <- read_csv("retrodiction/sims.csv")

for (i in 1:100) {
    print(i)
    stats <- data.frame(best = rep(0, 100 * 1001),
                    loglik = -Inf,
                    best0 = 0,
                    loglik0 = -Inf,
                    par = 0,
                    rep = 0,
                    strap = T)
    for (j in 1:100) {
        tmpstat <- suppressMessages(
            data.frame(t(read_csv(paste0("retrodiction/stats_", i, "_", j, ".csv"), col_names = F))))
        tmpstat$par <- i
        tmpstat$rep <- j
        tmpstat$strap <- T
        tmpstat$strap[1] <- F
        colnames(tmpstat) <- colnames(stats)
        idx <- 1001 * (j - 1)
        stats[(idx + 1):(idx + 1001), ] <- tmpstat
    }
    write_csv(stats, paste0("retrodiction/merge_", i, ".csv"))
}
stats <- data.frame(best = rep(0, 100 * 100 * 1001),
                    loglik = -Inf,
                    best0 = 0,
                    loglik0 = -Inf,
                    par = 0,
                    rep = 0,
                    strap = T)
for (i in 1:100) {
    print(i)
    tmpstat <- suppressMessages(read_csv(paste0("retrodiction/merge_", i, ".csv"), col_names = T))
    idx <- 100100 * (i - 1)
    stats[(idx + 1):(idx + 100100), ] <- tmpstat
}
write_csv(stats, "retrodiction/merged.csv")
stats <- read_csv("retrodiction/merged.csv")

stats <- cbind(stats, pars[stats$best, ])
bootstrapped <- stats %>%
    group_by(par, rep) %>% 
    summarize(estalpha = mean(alpha),
              estf = mean(f),
              estR0 = mean(R0))
bootstrapped <- cbind(bootstrapped, selected[rep(1:100, each = 100), 2:4])
bootstrapped %>% 
    summarize(lower = quantile(estR0, 0.025),
              mean = mean(estR0),
              upper = quantile(estR0, 0.975),
              true = mean(R0)) %>% 
    ggplot(aes(x = true, y = mean)) +
    geom_line() +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.1) +
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = 2) +
    theme_classic()
bootstrapped %>% 
    ggplot(aes(x = R0, y = estR0)) +
    geom_point(alpha = 0.05) +
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))

bootstrapped <- bootstrapped %>% 
    left_join(stats %>% 
                  filter(!strap) %>% 
                  select(par, rep, loglik, loglik0), 
              by = c("par", "rep")) %>% 
    mutate(emptest = pchisq(2*abs(loglik - loglik0), 
                            2, lower.tail = F, log.p = F) / 2)

sims$lambda <- apply(sims[, 1:50], 1, 
                     \(x) sum((x + 0.001) * (0:(O - 1))) / (N + 0.001 * O))
expected <- t(sapply(sims$lambda, \(x) N * dpois(0:(O - 1), x)))
sims$goodness <- pchisq(rowSums((sims[, 1:50] - expected) ^ 2 / expected), 
                   O - 2, lower.tail = F, log.p = F)

bootstrapped <- bootstrapped %>% 
    left_join(sims %>% select(par, rep, goodness), 
              by = c("par", "rep"))

bootstrapped %>% 
    mutate(is0 = ifelse(R0 == 0, "R0 = 0", "R0 > 0")) %>% 
    ggplot(aes(x = goodness)) +
    geom_histogram(color = "black", fill = "white") +
    xlab("Pearson's test") +
    facet_wrap(~is0) +
    geom_vline(xintercept = 0.01, color = "red", linetype = 2) +
    theme_classic()
bootstrapped %>% 
    mutate(is0 = ifelse(R0 == 0, "R0 = 0", "R0 > 0")) %>% 
    ggplot(aes(x = emptest)) +
    geom_histogram(color = "black", fill = "white") +
    xlab("p-value") +
    ggtitle("likelihood-ratio test") +
    facet_wrap(~is0) +
    geom_vline(xintercept = 0.01, color = "red", linetype = 2) +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))

```
