---
title: "fitting"
author: "Julian Gottfried"
date: "2026-07-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set-up
```{r}
library(tidyverse)
setwd("~/Desktop/CME")
```

## Local sim functions
```{r}
singlerun <- function(alphaS, alphaI, R0, G, S, I) {
    counter <- rep(0, 500)
    Ss <- rep(0, G)
    Is <- rep(0, G)
    t <- 0
    g <- 0
    while (g < G) {
        counter[S + I] = counter[S + I] + 1
        if (g >= t) {
            Ss[t + 1] <- S
            Is[t + 1] <- I
            t <- t + 1
        }
        if (S + I == 0) { foi = 0
        } else {foi <- R0 * S * I / (S + I)}
        rate <- alphaS + alphaI + foi + S + I
        g <- g - 1 / rate * log(1 - runif(1))
        pick <- runif(1)
        if (pick < alphaS / rate) { S <- S + 1
        } else if (pick < (alphaS + alphaI) / rate) { I <- I + 1
        } else if (pick < (alphaS + alphaI + foi) / rate) { 
            S <- S - 1
            I <- I + 1
        } else if (pick < (alphaS + alphaI + foi + S) / rate) { S <- S - 1
        } else {I <- I - 1}
    }
    c(Ss, Is)
}

plotsingle <- function(result, G) {
    data.frame(time = c(1:G, 1:G), size = result, state = c(rep("S", G), rep("I", G))) %>%
        mutate(state = factor(state, levels = c("S", "I"))) %>%
        ggplot(aes(x = time, y = size)) +
        geom_line() +
        facet_wrap(~state) +
        theme_classic() +
        theme(text=element_text(size=12,family="mono"),
              legend.position = "bottom")
}
```

## Grid parameters
```{r}
# alpha <- seq(from=0,by=1,to=99)
# f <- seq(from=0,by=0.005,to=0.495)
# R0 <- seq(from=0,by=0.1,to=9.9)

alphaS <- seq(from=0,by=1,to=99)
alphaI <- seq(from=0,by=0.1,to=9.9)
R0 <- seq(from=0,by=0.1,to=9.9)

N <- 42
O <- 50
```

## Aggregate cluster sims
```{r}
counts <- NULL
for (i in 1:100) {
    print(i)
    filename <- paste0("JLresults/counts_",i,".csv")
    if (file.exists(filename)) {
        countstmp <- data.frame(t(read.table(paste0("JLresults/counts_",i,".csv"), sep=",")))
    } else countstmp <- data.frame(matrix(NA, 100 ^ 2, O))
    countstmp$alphaS <- alphaS[i]
    countstmp$alphaI <- 0
    countstmp$R0 <- 0
    for (j in 1:100) {
        for (k in 1:100) {
            idx <- 100 * (j - 1) + k
            countstmp$alphaI[idx] <- alphaI[j]
            countstmp$R0[idx] <- R0[k]
        }
    }
    counts <- rbind(counts, countstmp)
}
write_csv(counts, "results2.csv")
```

## Convert counts
```{r}
counts <- read_csv("results.csv")
freqs <- counts
pseudo <- 1 / 10
freqs[, 1:O] <- freqs[, 1:O] + pseudo
rowsums <- rowSums(freqs[, 1:O])
freqs[, 1:O] <- freqs[, 1:O] %>% 
    mutate_all(\(x) x / rowsums)

freqs$alphaS[!is.na(freqs$X1)] %>% unique

write_csv(freqs, "pars.csv", col_names = T)
freqs <- read_csv("pars.csv", col_names = T)

freqs[14000, ] %>% 
    pivot_longer(starts_with("X"), names_to = "bin", values_to = "proportion") %>% 
    mutate(bin = as.numeric(str_remove(bin, "X")) - 1) %>%
    ggplot(aes(x = bin, y = proportion)) +
    geom_point()+
    theme_classic()

moments <- cbind(X1 = rowMeans(as.matrix(freqs[, 1:O]) %*% as.matrix(0:(O - 1))),
                 X2 = rowMeans(as.matrix(freqs[, 1:O]) %*% as.matrix((0:(O - 1)) ^ 2)),
                 freqs[, (O + 1):(O + 3)]) %>%
    mutate(mu = X1,
           var = X2 - X1 ^ 2) %>% 
    mutate(p = mu / var) %>% 
    mutate(r = mu * p / (1 - p))
moments %>% 
    filter(r > 0, p < 1, p > 0) %>% 
    mutate(f = alphaI / (alphaS + alphaI)) %>% 
    sample_n(10000) %>% 
    ggplot(aes(x = f, y = log(mu), color = alphaS)) + 
    geom_point(alpha = 0.1) +
    theme_classic()

cmemat %>% filter(c13) %>% pull(species) %>% unique %>% length

```

## Get CME data
```{r}
cmemat <- read.table("eitan_data/host_cme_matrix.txt", header = T, sep = "\t")
cmes <- colnames(cmemat %>% select(-host_id))
cmemat$subject <- str_split(cmemat$host_id, "_", simplify = T)[, 1]
cmemat$genome <- str_split(cmemat$host_id, "_", simplify = T)[, 2]
hostdat <-  read.table("eitan_data/hosts.txt", header = T, sep = "\t")
cmemat <- left_join(cmemat, hostdat %>% select(species, genus, family, host_id), by = "host_id")

cmetab <- read.table("eitan_data/cme_table.txt", header = T, sep = "\t")
cmetab %>% 
    ggplot(aes(x = host_count, y = species_count)) +
    geom_point(alpha = 0.5) +
    geom_abline(color = "red", intercept = 0, slope = 1, linetype = 2) +
    theme_classic()
cmetab %>% 
    arrange(host_count / species_count) %>% 
    filter(is_cme) %>% 
    mutate(cme = row_number()) %>% 
    ggplot(aes(x = cme, y = host_count / species_count)) +
    geom_point(alpha = 0.5) +
    geom_hline(yintercept = 1, color = "red", linetype = 2) +
    theme_classic()

cmedat <- cmemat %>% 
    group_by(subject, species) %>% 
    summarize_at(cmes, \(x) sum(x) != 0) %>% 
    select(-species) %>% 
    pivot_longer(cols = -subject, names_to = "cme") %>% 
    group_by(cme, subject) %>% 
    summarize(count = sum(value), .groups = "drop")

hostdat %>% group_by(aid, species) %>% summarize(n=n()) %>% pull(n) %>% table

alphas <- hostdat %>% 
    group_by(species) %>% summarize(alpha = n()) %>% mutate(alpha = alpha / n()) %>% 
    mutate(alpha = alpha * (hostdat %>% 
                              group_by(aid) %>% 
                              summarize(n = n()) %>% 
                              pull(n) %>% mean))
alphas1 <- hostdat %>% 
    group_by(species) %>% summarize(alpha = n()) %>% mutate(alpha = alpha / n()) %>% 
    mutate(alpha = alpha * (hostdat %>% 
                              group_by(aid) %>% 
                              summarize(n = n()) %>% 
                              pull(n) %>% mean))
alphas2 <- hostdat %>% 
    group_by(species) %>% summarize(alpha = n() / 42)

data.frame(x = alphas1$alpha, y = alphas2$alpha) %>% 
    ggplot(aes(x = x, y = y)) + 
    geom_point() + 
    theme_classic()

estf <- cmemat %>% 
    group_by(species, subject) %>% 
    summarize_at(cmes, \(x) sum(x) != 0) %>% 
    summarize_at(cmes, mean)
estf %>% 
    mutate(sums = rowSums(estf[, 2:length(cmes)])) %>% 
    left_join(alphas, by = "species") %>% 
    arrange(-sums * alpha) %>% 
    mutate(species = as.integer(factor(species, levels = unique(species)))) %>% 
    pivot_longer(all_of(cmes), names_to = "cme", values_to = "f") %>% 
    filter(f > 0) %>% 
    ggplot(aes(x = species, y = f * alpha, color = cme)) +
    geom_point(alpha = 0.5) +
    scale_color_viridis_d(option = 2) +
    guides(color = "none") +
    ylab("alphaI") +
    theme_classic()
estf %>% 
    mutate(sums = rowSums(estf[, 2:length(cmes)])) %>% 
    left_join(alphas, by = "species") %>% 
    arrange(-sums * alpha) %>% 
    mutate(species = as.integer(factor(species, levels = unique(species)))) %>% 
    pivot_longer(all_of(cmes), names_to = "cme", values_to = "f") %>% 
    filter(f > 0) %>% 
    ggplot(aes(x = alpha, y = f, color = species)) +
    geom_point(alpha = 0.5) +
    scale_color_viridis_c(option = 2) +
    guides(color = "none") +
    ylab("f") +
    theme_classic()

cmematsp <- cmemat %>% 
        group_by(subject, species) %>% 
        summarize_at(cmes, \(x) sum(x) != 0)
hostdatsp <- hostdat %>% 
        group_by(aid, species) %>% 
        summarize(n = n())
subjects <- unique(cmemat$subject)
pvals <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    cseg <- cmes[i]
    specieslist <- table(cmematsp[unlist(cmematsp[, cseg]), "species"])
    estp <- unlist(lapply(names(specieslist), \(x) 
                          specieslist[x] / table(hostdatsp$species)[x]))
    estp[names(estp)] <- mean(estp)
    pi <- rep(0, N)
    for (j in 1:N) {
        tester <- estp[hostdatsp[hostdatsp$aid == subjects[j], ]$species]
        length(tester[!is.na(tester)])
        pi[j] <- 1 - 
            prod(1 - estp[hostdatsp[hostdatsp$aid == subjects[j], ]$species], na.rm = T)
    }
    pvals[i] <- poisbinom::ppoisbinom(cmetab$subject_count[cmetab$cseg == cseg], 
                                      pi, lower_tail = T)
}
hist(pvals)

cmemat %>% filter(c4) %>% select(subject, species) %>% group_by(subject) %>% 
    distinct %>% pull(species) %>% length

# get alpha from data
store <- matrix(0, length(cmes), 2)
for (i in 1:length(cmes)) {
    fams <- cmemat %>% filter(.[[cmes[i]]]) %>% pull(family) %>% unique
    spstats <- cmemat %>% 
        filter(family %in% fams) %>% 
        group_by(subject) %>% 
        select(subject, species) %>%
        distinct %>% 
        summarize(n = n()) %>% 
        pull(n)
    store[i, ] <- c(mean(spstats), max(spstats))
}
hist(store[, 1], breaks = 20)
hist(round(store[, 1], 0), breaks = 20)
length(unique(round(store[, 1], 0)))

alpha <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    fams <- cmemat %>% filter(.[[cmes[i]]]) %>% pull(family) %>% unique
    alpha[i] <- cmemat %>% 
        filter(family %in% fams) %>% 
        group_by(subject) %>% 
        select(subject, species) %>%
        distinct %>% 
        summarize(n = n()) %>% 
        pull(n) %>% 
        mean
}
alpha <- data.frame(alpha = round(alpha, 0), cme = cmes)
```

## Pre-estimate f
```{r}
# Method that doesn't work
level = "genus"
estf <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    infecting <- cmemat[, level][cmemat[, cmes[i]]]
    estf[i] <- length(unique(infecting)) / length(infecting)
}
ggplot(data.frame(estf),
       aes(x = estf))+
    geom_histogram(color = "white", fill = "black") +
    theme_classic()

# Method that may work
coarse = "family"
fine = "species"
estf <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    infecting <- cmemat[, cmes[i]][cmemat[, fine] %in% 
                hostdat[, fine][hostdat[, coarse] %in% 
                                    unique(cmemat[, coarse][cmemat[, cmes[i]]])]]
    estf[i] <- sum(infecting)/length(infecting)
}
ggplot(data.frame(estf),
       aes(x = estf))+
    geom_histogram(color = "white", fill = "black") +
    xlab("f")+
    theme_classic()

# Snap to grid
closestf <- round(f[estf%/%0.005 + 1], 3)
ggplot(data.frame(closestf),
       aes(closestf))+
    geom_histogram(color = "white", fill = "black") +
    xlab("f")+
    theme_classic()+
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
```

## Do fitting
```{r}
pars <- slimbin %>% select(alphaS:R0)
fits <- data.frame(cme = cmes, idx = 0, loglik = -Inf)
for (i in 1:length(cmes)) {
    print(i)
    tmpdat <- cmedat %>% filter(cme == cmes[i]) %>% pull(count)
    tmpdat <- as.numeric(unname(table(factor(tmpdat, levels = 0:(O-1)))))
    factor <- log(factorial(N) / prod(factorial(tmpdat)))
    loglik <- rowSums(sweep(log(slimbin[, 1:O]), 2, tmpdat, FUN = "*"))
    fits[i, 2:3] <- c(which.max(loglik), factor + max(loglik))
}
write_csv(fits, "fits_slim.csv")
fits <- read_csv("fits_slim.csv")
```

## Do fitting without alpha
```{r}
pars <- freqs %>% select(alpha:R0)
fits <- data.frame(cme = cmes, idx = 0, loglik = -Inf)
for (i in 1:length(cmes)) {
    print(i)
    tmpalpha <- alpha$alpha[alpha$cme == cmes[i]]
    tmpdat <- cmedat %>% filter(cme == cmes[i]) %>% pull(count)
    tmpdat <- as.numeric(unname(table(factor(tmpdat, levels = 0:(O-1)))))
    factor <- log(factorial(N) / prod(factorial(tmpdat)))
    loglik <- rowSums(sweep(log(freqs[freqs$alpha == tmpalpha, 1:O]), 
                            2, tmpdat, FUN = "*"))
    fits[i, 2:3] <- c(which(freqs$alpha == tmpalpha)[1] + which.max(loglik) - 1,
                      factor + max(loglik))
}
write_csv(fits, "fits_slim.csv")
fits <- read_csv("fits_slim.csv")
```

## Do constrained fitting
```{r}
fits0 <- data.frame(cme = cmes, idx = 0, loglik = -Inf)
for (i in 1:length(cmes)) {
    print(i)
    tmpdat <- cmedat %>% filter(cme == cmes[i]) %>% pull(count)
    tmpdat <- as.numeric(unname(table(factor(tmpdat, levels = 0:(O-1)))))
    factor <- log(factorial(N) / prod(factorial(tmpdat)))
    loglik <- apply(slimbin[slimbin$R0 == 0, 1:O], 1, \(x) sum(tmpdat * log(x)))
    fits0[i, ] <- c(cmes[i], which.max(loglik), factor + max(loglik))
}
```

## Plot parameter spread
```{r}
explore <- pars[fits$idx, ] %>% mutate(cme = cmes)
explore$identified <- slimbin[fits$idx, ]$identifiable
sum(explore$identified)

explore %>% 
    arrange(-(alphaI + R0)) %>% 
    mutate(cme = row_number()) %>% 
    ggplot(aes(x = cme, y = R0 + alphaI)) +
    geom_hline(yintercept = 3, color = "red", linetype = 2) +
    geom_point(shape = 15, alpha = 0.5) +
    theme_classic() +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

explore %>% 
    group_by(alpha, f, R0) %>% 
    summarize(count = n()) %>% 
    ungroup %>% 
    pivot_longer(c(alpha,f,R0), names_to = "parameter") %>% 
    mutate(parameter = factor(parameter, 
                              levels = c("alpha","f","R0"))) %>% 
    ggplot(aes(x = value)) +
    geom_histogram(color = "white", fill = "black")+
    ggtitle("parameter fits across CMEs") +
    theme_classic() +
    facet_wrap(~parameter, scale="free") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

explore %>% 
    group_by(alphaS, alphaI, R0) %>%
    summarize(count = n()) %>%
    ungroup %>% 
    ggplot(aes(x = alphaS, y = R0)) +
    geom_point(aes(size = count), stroke = 0.5, alpha = 1, shape = 1) +
    scale_size(range = c(0.5, 10)) +
    ggtitle("parameter fits across CMEs") +
    theme_classic() +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

sum(explore$R0 != 0)
sum(explore$R0 == 0)
```

## Plot p fit
```{r}
ordering <- order(explore$f, decreasing = T)
data.frame(x = 1:length(cmes), 
           priorf = closestf[ordering],
           estf = explore$f[ordering]) %>% 
    pivot_longer(priorf:estf) %>% 
    ggplot(aes(x = x, y = value, color = name)) +
    scale_color_discrete(palette = c("#E0AB60", "#7DA2C9"), 
                         labels = c("estimated","calculated"),
                         name = "") +
    geom_point() +
    ylab("f") +
    xlab("CME") +
    theme_classic() +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
```

## Plot particular CMEs
```{r}
fits$cme[with(explore, R0 == max(R0))]
fits$cme[with(explore, R0 == 0)]
fits$cme[with(explore, alphaI == max(alphaI))]
fits$cme[with(explore, alphaI == min(alphaI))]

tester <- "c4"
cmedat %>%
    filter(cme == tester) %>%
    pull(count) %>% 
    factor(levels = 0:(O-1)) %>% 
    table %>% 
    data.frame %>% 
    rename(count = '.', freq = Freq) %>% 
    mutate(count = as.numeric(count) - 1, 
           freq = freq / sum(freq)) %>% 
    ggplot(aes(x=count, y=freq)) +
    geom_col(width = 1, color = "black", fill = "white")+
    ylim(0,1)+xlim(-1,51)+
    theme_classic() +
    ggtitle(tester) +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")


slimbin <- read_csv(paste0("alpharesults/alpha_",
                              which(alpha == min(fits$alpha[fits$cme == tester], 94)),
                              ".csv"))
round(N * slimbin[fits$idx[fits$cme == tester] + 100*50, 1:O], 1) %>% 
    pivot_longer(everything(), 
                 names_to = "count", 
                 values_to="freq") %>% 
    mutate(count=as.numeric(str_remove(count,"X"))-1, 
           freq = freq / sum(freq)) %>% 
    ggplot(aes(x=count, y=freq)) +
    geom_col(width = 1, color = "black", fill = "white")+
    ylim(0,1)+xlim(-1,51)+
    theme_classic() +
    ggtitle("best fit") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

testpars <- explore[explore$cme == tester, ]
#testpars$alphaS <- testpars$alpha * (1 - testpars$f)
#testpars$alphaI <- testpars$alpha * testpars$f
plotsingle(singlerun(testpars$alphaS,
                     testpars$alphaI, 
                     testpars$R0,
                     1000,
                     25, 0), 1000)
result <- singlerun(testpars$alphaS,
                     testpars$alphaI, 
                     testpars$R0,
                     1000,
                     25, 0)
Sres <- result[1:1000]
Ires <- result[1001:2000]
data.frame(t = 1:1000,freq = Ires/(Sres+Ires)) %>% 
    ggplot(aes(x = t, y = freq)) +
    geom_line() +
    theme_classic()
```

## Poisson goodness of fit
```{r}
goodness <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    tmpdat <- cmedat %>% filter(cme == cmes[i]) %>% pull(count)
    lambda <- mean(tmpdat)
    tmpdat <- as.numeric(unname(table(factor(tmpdat, levels = 0:(O-1)))))
    expected <- N*dpois(0:(O-1),lambda)
    stat <- sum((tmpdat-expected)^2 / expected)
    df <- O - 2
    goodness[i] <- pchisq(stat, df, lower.tail = F, log.p = T)
}

cmes[exp(goodness)<0.01]

data.frame(cme = cmes, pval = exp(goodness)) %>% 
    arrange(-pval) %>%
    mutate(idx = row_number()) %>% 
    ggplot(aes(x=idx,y=pval))+
    geom_point(alpha = 1, shape = 3)+
    geom_hline(yintercept = 0.01, color="red", linetype=2)+
    theme_classic()+
    xlab("cme")+ylab("p-value")+
    ggtitle("within-subject transmission") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
```

## Negative binomial error
```{r}
nsamp <- 10000
sampling <- freqs[sample(1:nrow(freqs), nsamp), ]
mse <- rep(0, nsamp)
for (i in 1:nsamp) {
    print(i)
    tmpdat <- as.numeric(sampling[i, 1:50])
    asint <- round(1000 * tmpdat, 0)
    p = mean(asint) / var(asint)
    r = mean(asint) * p / (1 - p)
    probs <- dnbinom(0:(O-1),r,p) / sum(dnbinom(0:(O-1),r,p))
    # dens[i] <- gtools::ddirichlet(tmpdat, probs)
    mse[i] <- mean((probs - tmpdat)^2)
}

data.frame(mse = mse) %>% 
    arrange(-mse) %>%
    mutate(sim = row_number()) %>% 
    ggplot(aes(x=sim,y=mse))+
    geom_line()+
    theme_classic()+
    ylim(0,1)+
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
```

## Negative binomial correspondence
```{r}
visit <- sample(1:nrow(pars), 10000)
abut <- cbind(pars[visit, ], mu = 0, sigmasq = 0, r = 0, p = 0)
for (i in 1:nrow(abut)) {
    if (i %% 100 == 1) print(i)
    tmpfreq <- as.numeric(freqs[visit, 1:50][i, ])
    mu <- sum(tmpfreq * (0:49))
    sigmasq <- sum(tmpfreq * (0:49) ^ 2) - mu ^ 2
    p = min(max(mu / sigmasq, 0), 1)
    r = max(floor(mu * p / (1 - p)), 0)
    abut[i, 4:7] <- c(mu, sigmasq, r, p)
}

abut %>% 
    filter(r < 50) %>% 
    ggplot(aes(x = alpha, y = f, fill = p)) +
    geom_tile() +
    theme_classic()
```

## Negative binomial likelihood ratio test
```{r}
lrt <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    tmpdat <- cmedat %>% filter(cme == cmes[i]) %>% pull(count)
    lambda <- mean(tmpdat)
    p = mean(tmpdat) / var(tmpdat)
    r = mean(tmpdat) * p / (1 - p)
    lr <- 2*abs(sum(dnbinom(tmpdat, r, p, log = T)) -
            sum(dpois(tmpdat, lambda, log = T)))
    df <- 1
    lrt[i] <- pchisq(lr, df, lower.tail = F, log.p = T) - log(2)
}

data.frame(cme = cmes, pval = exp(lrt)) %>% 
    arrange(-pval) %>%
    mutate(idx = row_number()) %>% 
    ggplot(aes(x=idx,y=pval))+
    geom_point(alpha = 1, shape = 3)+
    geom_hline(yintercept = 0.01, color="red", linetype=2)+
    theme_classic()+
    xlab("cme")+ylab("p-value")+
    ggtitle("within-subject transmission via LRT") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

data.frame(cme = cmes, good = exp(goodness), lrt = exp(lrt)) %>% 
    arrange(-good) %>%
    mutate(idx = row_number()) %>% 
    pivot_longer(good:lrt, names_to = "test", values_to = "pval") %>% 
    ggplot(aes(x=idx,y=pval, color=test))+
    geom_point(alpha = 0.5, shape = 3)+
    scale_color_discrete(palette=c("#c4933f","#2aa7f5"),
                         labels=c("goodness of fit", "LRT"),
                         name = "test")+
    geom_hline(yintercept = 0.01, color="red", linetype=2)+
    theme_classic()+
    xlab("cme")+ylab("p-value")+
    ggtitle("within-subject transmission") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
```

## Empirical likelihood ratio test
```{r}
emplrt <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    tmpdat <- cmedat %>% filter(cme == cmes[i]) %>% pull(count)
    lr <- 2*abs(as.numeric(fits[fits$cme == cmes[i], 3]) -
                    as.numeric(fits0[fits0$cme == cmes[i], 3]))
    df <- 2
    emplrt[i] <- pchisq(lr, df, lower.tail = F, log.p = T) - log(2)
}

data.frame(cme = cmes, pval = exp(emplrt), identified = explore$identified) %>% 
    filter(identified) %>% 
    arrange(-pval) %>%
    mutate(idx = row_number()) %>% 
    ggplot(aes(x=idx,y=pval))+
    geom_point(alpha = 1, shape = 3)+
    geom_hline(yintercept = 0.05, color="red", linetype=2)+
    theme_classic()+
    xlab("cme")+ylab("p-value")+
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

data.frame(cme = cmes, pval = exp(emplrt), identified = explore$identified) %>% 
    filter(identified, pval < 0.05) %>% pull(cme)

data.frame(cme = cmes, good = exp(goodness), lrt = exp(emplrt)) %>% 
    arrange(-good) %>%
    mutate(idx = row_number()) %>% 
    pivot_longer(good:lrt, names_to = "test", values_to = "pval") %>% 
    ggplot(aes(x=idx,y=pval, color=test))+
    geom_point(alpha = 0.5, shape = 3)+
    scale_color_discrete(palette=c("#c4933f","#2aa7f5"),
                         labels=c("goodness of fit", "empirical LRT"),
                         name = "test")+
    geom_hline(yintercept = 0.01, color="red", linetype=2)+
    theme_classic()+
    xlab("cme")+ylab("p-value")+
    ggtitle("within-subject transmission") +
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")
```

## alpha correlations
```{r}
taxa = c("species","genus","family")
prev <- matrix(0, length(cmes), 3, dimnames = list(NULL, taxa))
for (i in 1:length(cmes)) {
    infectable <- cmemat[, taxa][cmemat[, cmes[i]],]
    prev[i, ] <- unlist(lapply(taxa, \(taxon)
                               sum(table(hostdat[,taxon])[
                                   unique(infectable[,taxon])]
                                   ) / nrow(hostdat)))
}

explore %>% bind_cols(prev) %>% 
    pivot_longer(species:family, 
                 names_to="taxon", 
                 values_to="prevalence") %>% 
    ggplot(aes(x = prevalence, y = alpha))+
    geom_point() +
    theme_classic() +
    facet_wrap(~taxon, scales = "free_x")+
    xlab("prevalence of infectable taxa")+ylab("alpha")+
    theme(text=element_text(size=12,family="mono"),
          legend.position = "bottom")

# taxon = c("family")
# weighting <- matrix(0, 
#                length(cmes), 
#                length(unique(hostdat[, taxon])), 
#                dimnames = list(NULL, unique(hostdat[, taxon])))
# for (i in 1:length(cmes)) {
#     infectable <- unique(cmemat[, taxon][cmemat[, cmes[i]]])
#     tmpfam <- table(hostdat[, taxon])[infectable]
#     tmpfam <- tmpfam / sum(tmpfam)
#     weighting[i, names(tmpfam)] <- tmpfam * explore[i,]$R0
# }
# colSums(weighting)
```

## Matrix!
```{r}
ndim <- (O + 1) ^ 2

states <- matrix(0, ndim, 2)
k = 1
for (i in 0:O) {
    for (j in 0:O) {
        states[k, ] = c(i, j)
        k = k + 1
    }
}

Q <- matrix(0, ndim, ndim)

alpha1 = 25
alpha2 = 0.5
matR0 = 1

for (i in 1:ndim) {
    print(i)
    for (j in 1:ndim) {
        S = states[i, 1]
        I = states[i, 2]
        dS = states[j, 1] - S
        dI = states[j, 2] - I
        if (dS == 1 & dI == 0) {
            Q[i, j] = alpha1
        } else if (dS == 0 & dI == 1) {
            Q[i, j] = alpha2
        } else if (dS == -1 & dI == 0) {
            Q[i, j] = S
        } else if (dS == 0 & dI == -1) {
            Q[i, j] = I
        } else if (dS == -1 & dI == 1) {
            Q[i, j] = matR0 * S * I / (S + I)
        }
    }
}
for (i in 1:ndim) {
    Q[i, i] = -sum(Q[i, ])
}

pi <- pracma::nullspace(t(Q))
pi <- round(pi / sum(pi), 5)[, 1]

distribution <- data.frame(states) %>% 
    rename(S = X1, I = X2) %>% 
    mutate(pi = pi) %>% 
    group_by(I) %>% 
    summarize(pi = sum(pi))

ggplot(distribution, aes(x = I, y = pi)) +
    geom_point() + 
    theme_classic() +
    ylim(0, 1) +
    theme(text = element_text(size = 12, family = "mono"),
          legend.position = "bottom")
```

## UNHELPFUL retrodiction section
```{r}
selected <- read_csv("retrodiction/selected.csv", col_names = F)[1:100, ]
colnames(selected) <- c("idx", "alpha", "f", "R0")
selected$idx <- as.integer(selected$idx)

sims <- data.frame()
for (i in 1:100) {
    tmpsim <- data.frame(t(read_csv(paste0("retrodiction/dat_", i, ".csv"), col_names = F)))
    tmpsim$par <- i
    sims <- rbind(sims, tmpsim)
}
sims$rep <- rep(1:100, times = 100)
write_csv(sims, "retrodiction/sims.csv")
sims <- read_csv("retrodiction/sims.csv")

for (i in 1:100) {
    print(i)
    stats <- data.frame(best = rep(0, 100 * 1001),
                    loglik = -Inf,
                    best0 = 0,
                    loglik0 = -Inf,
                    par = 0,
                    rep = 0,
                    strap = T)
    for (j in 1:100) {
        tmpstat <- suppressMessages(
            data.frame(t(read_csv(paste0("retrodiction/stats_", i, "_", j, ".csv"), col_names = F))))
        tmpstat$par <- i
        tmpstat$rep <- j
        tmpstat$strap <- T
        tmpstat$strap[1] <- F
        colnames(tmpstat) <- colnames(stats)
        idx <- 1001 * (j - 1)
        stats[(idx + 1):(idx + 1001), ] <- tmpstat
    }
    write_csv(stats, paste0("retrodiction/merge_", i, ".csv"))
}
stats <- data.frame(best = rep(0, 100 * 100 * 1001),
                    loglik = -Inf,
                    best0 = 0,
                    loglik0 = -Inf,
                    par = 0,
                    rep = 0,
                    strap = T)
for (i in 1:100) {
    print(i)
    tmpstat <- suppressMessages(read_csv(paste0("retrodiction/merge_", i, ".csv"), col_names = T))
    idx <- 100100 * (i - 1)
    stats[(idx + 1):(idx + 100100), ] <- tmpstat
}
write_csv(stats, "retrodiction/merged.csv")
stats <- read_csv("retrodiction/merged.csv")

stats <- cbind(stats, pars[stats$best, ])
bootstrapped <- stats %>%
    group_by(par, rep) %>% 
    summarize(estalpha = mean(alpha),
              estf = mean(f),
              estR0 = mean(R0))
bootstrapped <- cbind(bootstrapped, selected[rep(1:100, each = 100), 2:4])
bootstrapped %>% 
    summarize(lower = quantile(estR0, 0.025),
              mean = mean(estR0),
              upper = quantile(estR0, 0.975),
              true = mean(R0)) %>% 
    ggplot(aes(x = true, y = mean)) +
    geom_line() +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = 0.1) +
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = 2) +
    theme_classic()
bootstrapped %>% 
    filter(R0 > 0) %>% 
    ggplot(aes(x = alpha*f*R0, y = estalpha*estf*estR0)) +
    geom_point(alpha = 0.05) +
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))

bootstrapped <- bootstrapped %>% 
    left_join(stats %>% 
                  filter(!strap) %>% 
                  select(par, rep, loglik, loglik0), 
              by = c("par", "rep")) %>% 
    mutate(emptest = pchisq(2*abs(loglik - loglik0), 
                            2, lower.tail = F, log.p = F) / 2)

sims$lambda <- apply(sims[, 1:50], 1, 
                     \(x) sum((x + 0.001) * (0:(O - 1))) / (N + 0.001 * O))
expected <- t(sapply(sims$lambda, \(x) N * dpois(0:(O - 1), x)))
sims$goodness <- pchisq(rowSums((sims[, 1:50] - expected) ^ 2 / expected), 
                   O - 2, lower.tail = F, log.p = F)

bootstrapped <- bootstrapped %>% 
    left_join(sims %>% select(par, rep, goodness), 
              by = c("par", "rep"))

bootstrapped %>% 
    mutate(is0 = ifelse(R0 == 0, "R0 = 0", "R0 > 0")) %>% 
    ggplot(aes(x = goodness)) +
    geom_histogram(color = "black", fill = "white") +
    xlab("Pearson's test") +
    facet_wrap(~is0) +
    geom_vline(xintercept = 0.01, color = "red", linetype = 2) +
    theme_classic()
bootstrapped %>% 
    mutate(is0 = ifelse(R0 == 0, "R0 = 0", "R0 > 0")) %>% 
    ggplot(aes(x = emptest)) +
    geom_histogram(color = "black", fill = "white") +
    xlab("p-value") +
    ggtitle("likelihood-ratio test") +
    facet_wrap(~is0) +
    geom_vline(xintercept = 0.01, color = "red", linetype = 2) +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))

stats %>%
    group_by(par, rep) %>% 
    mutate(og1 = alpha[1], og2 = f[1], og3 = R0[1]) %>%
    summarize(alpha = mean(alpha),
              f = mean(f),
              R0 = mean(R0),
              og1 = mean(og1),
              og2 = mean(og2),
              og3 = mean(og3)) %>%
    ggplot(aes(x = og1, y = alpha, color = og3)) +
    geom_point() +
    theme_classic()

# bootstrapping is generally correlated
# but i suppose why not just do exploration with both in turn

bootstrapped %>% 
    filter(R0>0) %>% 
    ggplot(aes(x = estR0, y = estalpha*estf, color = R0)) +
    geom_point(alpha = 0.1) +
    scale_color_viridis_c(option = 5) +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))

selected[1, ]
stats %>% filter(par == 100, rep == 15, strap) %>% 
    ggplot(aes(x = f)) +
    geom_density() +
    theme_classic()

```

## UNHELPFUL pca stuff
```{r}
dimred <- freqs[apply(freqs[, 1:O], 1, which.max) == 1, ]
sims <- matrix(0, O, nrow(dimred) * 10)
for (i in 1:nrow(dimred)) {
    if (i %% 1000 == 0) print(i)
    idx <- 10 * (i - 1)
    sims[, (idx + 1):(idx + 10)] <- rmultinom(10, N, dimred[i, 1:O])
}

# moments <- apply(sims, 2, \(f) {
#     x <- 0:(O - 1)
#     mu <- sum(x * f)
#     sigma <- sqrt(sum((x - mu) ^ 2 * f))
#     gamma <- sum(((x - mu) / sigma) ^ 3 * f)
#     c(mu, sigma ^ 2, gamma)
#     })

K <- 5
# pc <- prcomp(dimred[, 1:O], scale. = T, center = T)
pc <- prcomp(sims, scale. = T, center = T, rank. = K)
# pc <- prcomp(t(moments), scale. = T, center = T)

summary(pc)
rotated <- pc$x
loadings <- pc$rotation

cbind(rotated, dimred[, (O + 1):(O + 3)]) %>% 
    slice_sample(n = 10000) %>% 
    ggplot(aes(x = alpha, y = PC1)) +
    geom_point(alpha = 0.1) +
    theme_classic()

loadings %>% 
    data.frame %>% 
    mutate(idx = row_number()) %>% 
    pivot_longer(starts_with("PC")) %>% 
    ggplot(aes(x = idx, y = value, group = name, color = name)) +
    scale_color_viridis_d(direction = -1) +
    geom_line() +
    theme_classic()

M <- 100
J <- 1000
testidx <- sample(1:nrow(dimred), M, replace = F)
testers <- dimred[testidx, 1:O]
conv <- matrix(0, M, K)
for (i in 1:M) {
    print(i)
    sim <- as.numeric(rmultinom(1, N, testers[i, ]))
    simdat <- rep(0:(O - 1), sim)
    simfreq <- rep(0, O)
    for (j in 1:J) {
        simfreq <- simfreq + 
            as.numeric(table(factor(
                sample(simdat, N, replace = T), levels = 0:(O - 1)))) / 42
    }
    simfreq <- simfreq / J
    conv[i, ] <- ((simfreq - pc$center) / pc$scale) %*% loadings
}
mins <- matrix(0, M, nrow(dimred))
for (i in 1:nrow(dimred)) {
    if (i %% 1000 == 0) print(i)
    mins[, i] <- apply(conv, 1, \(x) sum(abs(rotated[i, ] - x)))
}
ests <- data.frame(idx = testidx, est = apply(mins, 1, which.min))

plot(dimred[ests$idx, (O + 1):(O + 3)]$alpha,
     dimred[ests$est, (O + 1):(O + 3)]$alpha)
plot(dimred[ests$idx, (O + 1):(O + 3)]$R0,
     dimred[ests$est, (O + 1):(O + 3)]$R0)
plot(dimred[ests$idx, (O + 1):(O + 3)]$f,
     dimred[ests$est, (O + 1):(O + 3)]$f)

sim <- as.numeric(rmultinom(1, N, testers[7, ]))

```

## Proper retrodiction
```{r}
lO <- 50
ok <- sapply(cmes, \(x) {
    cmedat %>%
        filter(cme == x) %>%
        pull(count) %>%
        factor(levels = 0:(O - 1)) %>%
        table %>%
        as.numeric
    }, simplify = T)
ok <- rbind(ok[1:(lO - 1), ], colSums(ok[lO:O, ]))
sum(apply(ok, 2, which.max) != 1)

binned <- cbind(freqs[, 1:(lO - 1)], 
                X = rowSums(freqs[, lO:O]),
                freqs[, (O+1):(O+3)])
colnames(binned)[lO] <- paste0("X", lO)
binned <- binned[!is.na(binned$X1), ]
slimbin <- binned[apply(binned[, 1:lO], 1, which.max) == 1, ]

slimbin %>% 
    mutate_at(paste0("X", 1:lO), \(x) round(x, 3)) %>%
    summarize(.by = starts_with("X"),
              alphaS = alphaS[1],
              alphaI = alphaI[1],
              R0 = R0[1],
              identifiable = n() == 1) %>% 
    ggplot(aes(x = R0, y = alphaI, fill = identifiable)) + 
    geom_tile(aes(alpha = !identifiable), color = "black", linewidth = 0.01) +
    scale_fill_manual(values = c("black","white"),
                      labels = c("non-identifiable", "identifiable"),
                      name = "") +
    guides(alpha = "none") +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))

slimbin <- slimbin %>% 
    summarize(.by = starts_with("X"),
              alpha = alpha[1],
              f = f[1],
              R0 = R0[1],
              identifiable = n() == 1)

slimbin[, 1:lO] <- t(apply(slimbin[, 1:lO], 1, \(x) (x + 0.00001) / sum(x + 0.00001)))

logged <- as.matrix(log(slimbin[, 1:lO]))
# logged0 <- logged[slimbin$R0 == 0, ]

npars <- 200
M <- 1
J <- 100
testidx <- sample((1:nrow(slimbin))[with(slimbin, R0 <= 5 & f <= 0.5)], npars, replace = F)
# sum(slimbin[testidx, ]$R0 == 0)
# testidx <- c(sample((1:nrow(slimbin))[slimbin$R0 == 0], npars / 2, replace = F),
#              sample((1:nrow(slimbin))[slimbin$R0 != 0], npars / 2, replace = F))
# testidx <- unlist(sapply(seq(from = 0, by = 0.1, to = 2.9), \(x) {
#     chunk <- (1:nrow(slimbin))[slimbin$R0 == 1 & slimbin$alphaI == x]
#     if (length(chunk) == 0) {
#         returner <- NULL
#     } else returner <- sample((1:nrow(slimbin))[slimbin$R0 == 1 & slimbin$alphaI == x], 
#            7, replace = F)
#     returner}))
# testidx <- unlist(sapply(seq(from = 0, by = 0.1, to = 2.9), \(x) {
#     chunk <- (1:nrow(slimbin))[slimbin$R0 == x & slimbin$alphaI == 1]
#     if (length(chunk) == 0) {
#         returner <- NULL
#     } else returner <- sample((1:nrow(slimbin))[slimbin$R0 == x & slimbin$alphaI == 1], 
#            7, replace = F)
#     returner}))
# npars <- length(testidx)

bestidx <- matrix(0, M * J, npars)
emplrt <- matrix(0, M * J, npars)
# lrt <- matrix(0, M * J, npars)
# goodness <- matrix(0, M * J, npars)
for (i in 1:npars) {
    print(i)
    # i <- 100
    tmpalpha <- slimbin$alpha[testidx[i]]
    tmplogged <- logged[slimbin$alpha == tmpalpha, ]
    tmplogged0 <- logged[slimbin$alpha == tmpalpha & slimbin$R0 == 0, ]

    tmpdat <- rmultinom(M, N, slimbin[testidx[i], 1:lO])
    tmpdatresample <- matrix(0, lO, M * J)
    if (J == 1) { tmpdatresample <- tmpdat
    } else {
        for (j in 1:J) {
            tmpdatresample[, M * (j - 1) + 1:M] <- apply(tmpdat, 2, \(x) {
                table(factor(sample(rep(0:(lO - 1), as.numeric(x)), N, replace = T), 
                             levels = 0:(lO - 1)))
            })
        }
    }
    bestidx[, i] <- apply(tmpdatresample, 2, \(x) which.max(tmplogged %*% x)) +
        which(slimbin$alpha == tmpalpha)[1] - 1
    
    best <- apply(tmpdatresample, 2, \(x) max(tmplogged %*% x))
    
    best0 <- apply(tmpdatresample, 2, \(x) max(tmplogged0 %*% x))
    emplrt[, i] <- pchisq(2 * abs(best - best0), 1, lower.tail = F, log.p = F) / 2

    # pois <- apply(tmpdatresample, 2, \(x) {
    #     lambda <- mean(rep(0:(lO - 1), as.numeric(x))) + 1e-5
    #     probs <- dpois(0:(lO-1), lambda)
    #     sum(log(probs) * x)
    # })
    # lrt[, i] <- pchisq(2 * abs(best - pois), 2, lower.tail = F, log.p = F) / 2
    # 
    # expected <- apply(tmpdatresample, 2, \(x) {
    #     lambda <- mean(rep(0:(lO - 1), as.numeric(x)))
    #     exp <- N * dpois(0:(lO-1), lambda) + 0.0001
    #     42 * exp / sum(exp)
    # })
    # goodness[, i] <- pchisq(colSums((tmpdatresample - expected) ^ 2 / expected), 
    #                       lO - 2, lower.tail = F, log.p = F)
}
# write_csv(data.frame(testidx), "holding/testidx6.csv")
# write_csv(data.frame(bestidx), "holding/bestidx6.csv")
# write_csv(data.frame(emplrt), "holding/emplrt6.csv")

# testidx <- as.matrix(read_csv("holding/testidx2.csv"))[, 1]
# bestidx <- as.matrix(read_csv("holding/bestidx2.csv"))
# emplrt <- as.matrix(read_csv("holding/emplrt2.csv"))

retro <- matrix(0, 4 * npars, 4)
for (i in 1:npars) {
    retro[4 * (i - 1) + 1:4, ] <- 
        as.matrix(rbind(slimbin[testidx[i], (lO + 1):(lO + 4)],
                        cbind(apply(slimbin[bestidx[, i], (lO + 1):(lO + 3)], 2,
                                \(x) c(quantile(x, 0.1), 
                                       quantile(x, 0.5), 
                                       quantile(x, 0.9))),
                            identifiable = mean(slimbin[bestidx[, i], ]$identifiable))))
}
retro <- data.frame(cbind(retro[rep(c(T, F, F, F), npars), 1:3],
                         retro[rep(c(F, T, F, F), npars), 1:3],
                         retro[rep(c(F, F, T, F), npars), 1:3],
                         retro[rep(c(F, F, F, T), npars), 1:3],
                         retro[rep(c(T, F, F, F), npars), 4],
                         retro[rep(c(F, T, F, F), npars), 4]))
retro <- retro %>% 
    rename(alpha = X1, f = X2, R0 = X3,
           alphalow = X4, flow = X5, R0low = X6,
           alphamed = X7, fmed = X8, R0med = X9,
           alphahigh = X10, fhigh = X11, R0high = X12,
           identifiable = X13, identified = X14)

# plotretro <- rbind(retro[sample((1:nrow(retro))[retro$R0 == 0], 10, replace = F), ],
#       retro[retro$R0 != 0, ]) %>% 
#     filter(identified == 1)
plotretro <- retro %>% filter(identified == 1)
plotretro <- retro

ggplot(plotretro, aes(x = alpha, y = alphamed)) +
    geom_point(alpha = 0.6, shape = 12) +
    geom_linerange(aes(ymin = alphalow, ymax = alphahigh),
                    alpha = 0.15) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))
ggplot(plotretro, aes(x = f, y = fmed)) +
    geom_linerange(aes(ymin = flow, ymax = fhigh),
                    alpha = 0.15) +
    geom_point(alpha = 0.6, shape = 12) +
    scale_color_manual(values = c("grey", "black")) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))
ggplot(plotretro, aes(x = R0, y = R0med)) +
    geom_linerange(aes(ymin = R0low, ymax = R0high),
                    alpha = 0.15) +
    geom_point(alpha = 0.6, shape = 12) +
    scale_color_manual(values = c("grey", "black")) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))
ggplot(plotretro, aes(x = R0 + alpha*f, y = R0med + alphamed*fmed)) +
    geom_linerange(aes(ymin = R0low + alphalow*flow, ymax = R0high + alphahigh*fhigh),
                    alpha = 0.15) +
    geom_point(alpha = 0.6, shape = 12) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))

ggplot(plotretro, aes(x = alphaI, y = alphaImed, color = R0 < 3)) +
    geom_linerange(aes(ymin = alphaIlow, ymax = alphaIhigh),
                    alpha = 0.15) +
    geom_point(alpha = 0.6, shape = 12) +
    scale_color_manual(values = c("grey", "black")) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))
ggplot(plotretro, aes(x = R0, y = R0med, color = alphaI < 3)) +
    geom_linerange(aes(ymin = R0low, ymax = R0high),
                    alpha = 0.15) +
    geom_point(alpha = 0.6, shape = 12) +
    scale_color_manual(values = c("grey", "black")) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))

plotretro %>% filter(R0med + alphaImed < 3) %>% 
    ggplot(aes(x = alphaI, y = alphaImed)) +
    geom_linerange(aes(ymin = alphaIlow, ymax = alphaIhigh),
                    alpha = 0.15) +
    geom_point(alpha = 0.6, shape = 12) +
    scale_color_manual(values = c("grey", "black")) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))
plotretro %>% filter(R0med + alphaImed < 3) %>% 
    ggplot(aes(x = R0, y = R0med)) +
    geom_linerange(aes(ymin = R0low, ymax = R0high),
                    alpha = 0.15) +
    geom_point(alpha = 0.6, shape = 12) +
    scale_color_manual(values = c("grey", "black")) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))
plotretro %>% filter(R0med + alphaImed < 3) %>% 
    ggplot(aes(x = R0 + alphaI, y = R0med + alphaImed)) +
    geom_linerange(aes(ymin = R0low + alphaIlow, ymax = R0high + alphaIhigh),
                    alpha = 0.15) +
    geom_point(alpha = 0.6, shape = 12) +
    scale_color_manual(values = c("grey", "black")) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))

# plotretro %>% 
#     filter(alphaImed + R0med < 3) %>% 
#     ggplot(aes(x = R0med, y = alphaImed, color = R0)) +
#     geom_point(alpha = 0.9, shape = 12) +
#     geom_abline(slope = -1, intercept = seq(from = 0, to = 7, by = 1/2), alpha = 0.25) +
#     theme_classic() +
#     theme(text = element_text(size = 12, family = "mono"))

data.frame(t(emplrt)) %>% 
    mutate(is0 = ifelse(retro$R0 == 0, "R0 = 0", "R0 > 0"),
           identified = retro$identified) %>% 
    filter(identified == 1) %>% 
    select(-identified) %>% 
    pivot_longer(-is0) %>% 
    ggplot(aes(x = value)) +
    geom_histogram(color = "black", fill = "white") + 
    facet_wrap(~is0) +
    xlab("p-value") +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))
```

## Proper retrodictive significance testing
```{r}
npars <- 900
M <- 2
J <- 1
testidx <- c(sample((1:nrow(slimbin))[slimbin$R0 == 0], npars / 2, replace = F),
             sample((1:nrow(slimbin))[slimbin$R0 != 0], npars / 2, replace = F))
emplrt <- matrix(0, M * J, npars)
lrt <- matrix(0, M * J, npars)
goodness <- matrix(0, M * J, npars)
for (i in 1:npars) {
    print(i)
    tmpdat <- rmultinom(M, N, slimbin[testidx[i], 1:lO])
    tmpdatresample <- matrix(0, lO, M * J)
    if (J == 1) { tmpdatresample <- tmpdat
    } else {
        for (j in 1:J) {
            tmpdatresample[, M * (j - 1) + 1:M] <- apply(tmpdat, 2, \(x) {
                table(factor(sample(rep(0:(lO - 1), as.numeric(x)), N, replace = T), 
                             levels = 0:(lO - 1)))
            })
        }
    }
    
    best <- apply(tmpdatresample, 2, \(x) max(logged %*% x))
    
    pois <- apply(tmpdatresample, 2, \(x) {
        lambda <- mean(rep(0:(lO - 1), as.numeric(x)))
        probs <- dpois(0:(lO-1), lambda) + 1e-5
        sum(log(probs) * x)
    })
    best0 <- apply(tmpdatresample, 2, \(x) max(logged0 %*% x))
    
    lrt[, i] <- pchisq(2 * abs(best - pois), 2, lower.tail = F, log.p = F) / 2
    emplrt[, i] <- pchisq(2 * abs(best - best0), 2, lower.tail = F, log.p = F) / 2
    
    expected <- apply(tmpdatresample, 2, \(x) {
        lambda <- mean(rep(0:(lO - 1), as.numeric(x)))
        exp <- N * dpois(0:(lO-1), lambda) + 0.0001
        42 * exp / sum(exp)
    })
    goodness[, i] <- pchisq(colSums((tmpdatresample - expected) ^ 2 / expected), 
                          lO - 2, lower.tail = F, log.p = F)
}

data.frame(t(lrt)) %>% 
    mutate(is0 = c(rep("R0 = 0", npars / 2),
                   rep("R0 > 0", npars / 2)),
           R0 = slimbin$R0[testidx]) %>% 
    pivot_longer(-c(is0, R0)) %>% 
    ggplot(aes(x = value)) +
    geom_histogram() + 
    facet_wrap(~is0) +
    theme_classic()
data.frame(t(emplrt)) %>% 
    mutate(is0 = c(rep("R0 = 0", npars / 2),
                   rep("R0 > 0", npars / 2)),
           R0 = slimbin$R0[testidx]) %>% 
    pivot_longer(-c(is0, R0)) %>% 
    ggplot(aes(x = value)) +
    geom_histogram() + 
    facet_wrap(~is0) +
    theme_classic()
data.frame(t(goodness)) %>% 
    mutate(is0 = c(rep("R0 = 0", npars / 2),
                   rep("R0 > 0", npars / 2)),
           R0 = slimbin$R0[testidx]) %>% 
    pivot_longer(-c(is0, R0)) %>% 
    ggplot(aes(x = value)) +
    geom_histogram() + 
    facet_wrap(~is0) +
    theme_classic()

data.frame(t(emplrt)) %>% 
    mutate(is0 = c(rep("R0 = 0", npars / 2),
                   rep("R0 > 0", npars / 2)),
           R0 = slimbin$R0[testidx]) %>% 
    pivot_longer(-c(is0, R0)) %>% 
    group_by(R0) %>% 
    summarize(call = mean(value < 0.05)) %>% 
    ggplot(aes(x = R0, y = call)) +
    geom_point() + 
    theme_classic()

freqs[sample((1:nrow(freqs))[!is.na(freqs$X1)], 1000, replace = F), (O+1):(O+3)] %>% pairs
```

```{r}
taxon <- "family"
tmphost <- hostdat %>%
    mutate(subject = str_split(host_id, "_", simplify = T)[, 1]) %>% 
    group_by(subject, .data[[taxon]]) %>% 
    summarize(n = n(), family = family[1])
tmphost %>% 
    summarize(N = sum(n)) %>% 
    right_join(tmphost, by = "subject") %>% 
    mutate(freq = n / N) %>% 
    group_by(.data[[taxon]]) %>% 
    mutate(avg = mean(freq)) %>% 
    ungroup %>% 
    arrange(freq) %>% 
    mutate(idx = row_number()) %>% 
    mutate(taxon = factor(.data[[taxon]], levels = unique(.data[[taxon]]))) %>% 
    ggplot(aes(x = idx, y = freq)) +
    geom_point(size = 0.5, alpha = 0.25) +
    theme_classic() +
    xlab(taxon) + ylab("frequency") +
    theme(axis.text.x = element_blank(),
          axis.ticks.x = element_blank()) +
    theme(text = element_text(size = 12, family = "mono"))

```