---
title: "fitting2"
author: "Julian Gottfried"
date: "2026-08-21"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setup
```{r}
library(tidyverse)
setwd("~/Desktop/CME")
```

## Functions
```{r}
singlerun <- function(alpha, f, R0, G, S, I) {
    Ss <- matrix(0, nrow = 168, ncol = G)
    Is <- matrix(0, nrow = 168, ncol = G)
    t <- rep(0, 168)
    g <- 0
    
    while (g < G) {
        pop = S + I
        if (pop > 0 & pop <= 168) {
            if (g >= t[pop]) {
                t[pop] <- t[pop] + 1
                Ss[pop, t[pop]] <- S
                Is[pop, t[pop]] <- I
            }
        }
        if (S + I == 0) { foi = 0
        } else {foi <- R0 * S * I / (S + I)}
        rate <- alpha + foi + S + I
        g <- g - 1 / rate * log(1 - runif(1))
        pick <- runif(1)
        if (pick < alpha * (1 - f) / rate) { S <- S + 1
        } else if (pick < alpha / rate) { I <- I + 1
        } else if (pick < (alpha + foi) / rate) { 
            S <- S - 1
            I <- I + 1
        } else if (pick < (alpha + foi + S) / rate) { S <- S - 1
        } else {I <- I - 1}
    }
    list(Ss, Is)
}

fitting <- function(statistics, counts) {
    factor <- 0
    loglik <- rep(0, 100*100)
    for (j in unique(statistics$N)) {
        if (j == 0) next
        slimdata <- statistics %>% filter(N == j)
        slimI <- as.numeric(unname(table(factor(slimdata$I, levels = 0:(O-1)))))
        factor <- factor + log(factorial(sum(slimI))) - sum(log(factorial(slimI)))
        slimfreq <- counts %>% filter(N == j)
        loglik <- loglik + apply(slimfreq[, 1:O], 1, \(x) sum(slimI * log(x)))
    }
    
    # https://journals.sagepub.com/doi/pdf/10.1177/1536867X0700700305
    
    hat <- which.max(loglik)
    
    profiledf <- sapply(f, \(x) max(loglik[round(slimfreq$f, 2) == round(x, 2)]))
    finterval <- 2 * (max(loglik) - profiledf) <= qchisq(0.89, 1)

    profiledR0 <- sapply(R0, \(x) max(loglik[round(slimfreq$R0, 2) == round(x, 2)]))
    R0interval <- 2 * (max(loglik) - profiledR0) <= qchisq(0.89, 1)

    returner <- rep(0, 8)
    
    returner[1] <- slimfreq$f[hat]
    returner[2] <- range(f[finterval])[1]
    returner[3] <- range(f[finterval])[2]

    returner[4] <- slimfreq$R0[hat]
    returner[5] <- range(R0[R0interval])[1]
    returner[6] <- range(R0[R0interval])[2]

    returner[7] <- pchisq(2 * (max(loglik) - max(loglik[slimfreq$R0 == 0])), 
                            1, lower.tail = F, log.p = F) / 2
    
    returner[8] <- max(loglik)
    
    returner
}
```

## Hyperparams
```{r}
N <- 42
O <- 50

f <- seq(from = 0, to = 0.99, by = 0.01)
R0 <- seq(from = 0, to = 9.9, by = 0.1)
```

## CME data
```{r}
cmemat <- read.table("eitan_data/host_cme_matrix.txt", header = T, sep = "\t")
cmes <- colnames(cmemat %>% select(-host_id))
cmemat$aid <- str_split(cmemat$host_id, "_", simplify = T)[, 1]
cmemat$genome <- str_split(cmemat$host_id, "_", simplify = T)[, 2]
hostdat <-  read.table("eitan_data/hosts.txt", header = T, sep = "\t")
cmemat <- left_join(cmemat, hostdat %>% select(species, genus, family, host_id), by = "host_id")
cmetab <- read.table("eitan_data/cme_table.txt", header = T, sep = "\t")
cmedat <- cmemat %>% 
    group_by(aid, species) %>% 
    summarize_at(cmes, \(x) sum(x) != 0) %>% 
    select(-species) %>% 
    pivot_longer(cols = -aid, names_to = "cme") %>% 
    group_by(cme, aid) %>% 
    summarize(count = sum(value), .groups = "drop")
```

## N and I data
```{r}
NI <- data.frame(matrix(0, N * length(cmes), 3))
colnames(NI) <- c("cme","N","I")
for (i in 1:length(cmes)) {
    print(i)
    fams <- cmemat %>% filter(.[[cmes[i]]]) %>% pull(family) %>% unique
    Nstats <- suppressMessages(cmemat %>% 
        filter(family %in% fams) %>% 
        group_by(aid) %>% 
        select(aid, species) %>%
        distinct %>% 
        summarize(N = n()))
    Istats <- suppressMessages(cmemat %>% filter(.[[cmes[i]]]) %>% group_by(aid, species) %>% 
        summarize(n = n()) %>% summarize(I = n()))
    statistics <- left_join(Nstats, Istats, by = "aid")
    statistics <- rbind(statistics,
                        data.frame(aid = unique(hostdat$aid), N = 0, I = NA)[!(unique(hostdat$aid) %in% statistics$aid), ])
    statistics[is.na(statistics)] <- 0
    NI[N * (i - 1) + (1:N), ] <- statistics %>% mutate(cme = cmes[i]) %>% select(cme, N, I)
}
```

## Estimate alpha
```{r}
alphas <- NI %>% mutate(cme = factor(NI$cme, levels = unique(NI$cme))) %>% 
    group_by(cme) %>% summarize(alpha = round(mean(N, 0)))
alphalist <- sort(unique(alphas$alpha))
```

## Retrodiction: setup
```{r}
testalpha <- 33
testcounts <- read_csv(paste0("alpharesults/alpha_", which(alphalist == testalpha),".csv"))
testcounts[, 1:O] <- testcounts[, 1:O] + 1 / 10
rowsums <- rowSums(testcounts[, 1:O])
testcounts[, 1:O] <- testcounts[, 1:O] %>% 
    mutate_all(\(x) x / rowsums)

condN <- rpois(N, testalpha)
samplepars <- rbind(testcounts %>% filter(R0 == 0, f <= 0.5) %>% 
                        select(f, R0) %>% sample_n(100),
                    testcounts %>% filter(R0 > 0, R0 < 5, f <= 0.5) %>% 
                        select(f, R0) %>% sample_n(100))
```

## Retrodiction: fitting
```{r}
testret <- matrix(0, nrow=200, ncol=9)
for (i in 1:200) {
    print(i)
    sim <- apply((testcounts %>% filter(f == samplepars[i, ]$f, R0 == samplepars[i, ]$R0))[condN, 1:O],
                 1, \(x)  which.max(rmultinom(1, 1, x)) - 1)
    
    statistics <- data.frame(N = condN, I = sim)
    
    testret[i, 1:8] <- fitting(statistics, testcounts)
    
    fj <- statistics %>% filter(N > 0) %>% 
        mutate(f = I / N) %>% pull(f) %>% mean
    pi <- 1 - (1 - fj) ^ statistics$N
    if (sum(pi) == 0) { testret[i, 9] <- 1
    } else {
        obs <- sum(statistics$I > 0)
        testret[i, 9] <- poisbinom::ppoisbinom(obs, pi)
    }
}
```

## Retrodiction: viz
```{r}
explore <- data.frame(cbind(testret, samplepars))
colnames(explore)[1:9] <- c("fhat","fl","fu","R0hat","R0l","R0u","pval4","loglik","pval3")

explore %>% 
    ggplot(aes(x = R0, y = (R0l + R0u) / 2)) +
    geom_linerange(aes(ymin = R0l, ymax = R0u), alpha = 0.15) +
    geom_point(alpha = 0.9, shape = 1) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))

explore %>% 
    ggplot(aes(x = f, y = (fl + fu) / 2)) +
    geom_linerange(aes(ymin = fl, ymax = fu), alpha = 0.15) +
    geom_point(alpha = 0.9, shape = 1) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic() +
    ylab("estimate") +
    theme(text = element_text(size = 12, family = "mono"))

explore %>%
    mutate(is0 = ifelse(R0 == 0, "R0 = 0", "R0 > 0")) %>% 
    ggplot(aes(x = pval4)) +
    geom_histogram(color = "black", fill = "white") + 
    facet_wrap(~is0) +
    xlab("p-value") +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))

explore %>%
    mutate(is0 = ifelse(R0 == 0, "R0 = 0", "R0 > 0")) %>% 
    ggplot(aes(x = pval3)) +
    geom_histogram(color = "black", fill = "white") + 
    facet_wrap(~is0) +
    xlab("p-value") +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))
```

## CME fitting
```{r}
fits <- data.frame(cme = cmes, 
                   fhat = 0, fl = 0, fu = 0,
                   R0hat = 0, R0l = 0, R0u = 0,
                   pval = 1,
                   loglik = -Inf)
idx <- 1
for (i in 1:45) {
    tmpalpha <- alphalist[i]
    counts <- suppressMessages(read_csv(paste0("alpharesults/alpha_", i, ".csv")))
    counts[, 1:O] <- counts[, 1:O] + 1 / 10
    divider <- rowSums(counts[, 1:O])
    counts[, 1:O] <- apply(counts[, 1:O], 2, \(x) x / divider)
    
    if (i == 45) { tmpcmes <- alphas$cme[alphas$alpha >= tmpalpha]
    } else tmpcmes <- alphas$cme[alphas$alpha == tmpalpha]
    
    for (j in 1:length(tmpcmes)) {
        print(idx)
        tmpNI <- NI %>% filter(cme == tmpcmes[j])
        fits[idx, ] <- c(tmpcmes[j], fitting(tmpNI, counts))
        idx <- idx + 1
    }
}
write_csv(fits, "newfits2.csv")
```

## Viz
```{r}
fits <- fits %>% 
    mutate(across(-cme, as.numeric))

NI %>% group_by(cme) %>% summarize(test = sum(I)/sum(N)) %>% 
    right_join(fits, by = "cme") %>% 
    mutate(failed = R0u == 9.9 & R0l == 0,
           test = round(test, 3)) %>% 
    group_by(test) %>% summarize(pfail = mean(failed)) %>% 
    ggplot(aes(x = test, y = pfail)) +
    geom_point(size = 2.5, shape = 10) +
    theme_classic() +
    xlab("fraction of observations infected") + ylab("fitting failure") +
    theme(text = element_text(size = 12, family = "mono"))

fits %>% 
    arrange(-(R0l + R0u) / 2) %>% 
    mutate(cme = row_number(), sig = pval >= 0.05) %>% 
    ggplot(aes(x = cme, y = (R0l + R0u) / 2, color = sig)) +
    geom_linerange(aes(ymin = R0l, ymax = R0u), alpha = 0.15) +
    geom_point(alpha = 0.8, shape = 3) +
    scale_color_manual(name = "",
                       labels = c("significant","non-significant"),
                       values = c("#910027", "#818ade")) +
    theme_classic() +
    ylab("R0") +
    theme(text = element_text(size = 12, family = "mono"))

fits %>% 
    arrange(-(fl + fu) / 2) %>% 
    mutate(cme = row_number()) %>% 
    ggplot(aes(x = cme, y = (fl + fu) / 2)) +
    geom_linerange(aes(ymin = fl, ymax = fu), alpha = 0.15) +
    geom_point(alpha = 0.8, shape = 3) +
    theme_classic() +
    ylab("f") +
    theme(text = element_text(size = 12, family = "mono"))

fits %>% 
    arrange(-pval) %>% 
    mutate(cme = row_number()) %>% 
    ggplot(aes(x = cme, y = pval)) +
    geom_point(alpha = 0.8, shape = 3) +
    geom_hline(yintercept = 0.05, color = "red", linetype = 2) +
    theme_classic() +
    ylab("p-value") +
    theme(text = element_text(size = 12, family = "mono"))
```

## Examining specific CMEs
```{r}
tester <- "c8"
parfit <- fits[fits$cme == tester, c("fl", "fu", "R0l", "R0u")]
testcounts <- read_csv(paste0("alpharesults/alpha_",
                          which(alphalist == min(alphas$alpha[alphas$cme == tester], 94)),
                          ".csv"))
testcounts <- testcounts %>% filter(f == with(parfit, round((fl + fu) / 2, 2)), 
                                    R0 == with(parfit, round((R0l + R0u) / 2, 1)))
gatherNI <- colSums(testcounts[NI[NI$cme == tester, "N"], ])
gatherNI <- gatherNI / sum(gatherNI)

cmedat %>%
    filter(cme == tester) %>%
    pull(count) %>% 
    factor(levels = 0:(O-1)) %>% 
    table %>% 
    data.frame %>% 
    rename(count = '.', freq = Freq) %>% 
    mutate(count = as.numeric(count) - 1, 
           freq = freq / sum(freq)) %>% 
    ggplot(aes(x = count, y = freq)) +
    geom_col(width = 1, color = "black", fill = "white")+
    ylim(0, 1) + xlim(-1, 51) +
    theme_classic() +
    ggtitle(tester) +
    theme(text = element_text(size = 12, family = "mono"),
          legend.position = "bottom")

data.frame(count = 0:(O - 1), freq = gatherNI[1:O]) %>% 
    ggplot(aes(x = count, y = freq)) +
    geom_col(width = 1, color = "black", fill = "white")+
    ylim(0, 1) + xlim(-1, 51) +
    theme_classic() +
    ggtitle("best fit") +
    theme(text = element_text(size = 12, family = "mono"),
          legend.position = "bottom")
```

## Permutation --> Model: step 1
```{r}
pval1 <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    fj <- suppressMessages(cmemat %>% mutate(inq = cmemat[[cmes[i]]]) %>% 
        group_by(species) %>% summarize(f = mean(inq)))
    pi <- left_join(hostdat, fj, by = "species") %>%
        group_by(aid) %>%
        select(aid, species, f) %>% 
        distinct %>% 
        summarize(pi = 1 - prod(1 - f)) %>%
        pull(pi)
    obs <- cmemat %>% mutate(inq = cmemat[[cmes[i]]]) %>% 
        group_by(aid) %>% summarize(present = sum(inq) > 0) %>% 
        pull(present) %>% sum
    pval1[i] <- poisbinom::ppoisbinom(obs, pi)
}
data.frame(cme = 1:length(cmes), pval = sort(pval1, decreasing = T)) %>% 
    ggplot(aes(x = cme, y = pval)) +
    geom_point(alpha = 0.8, shape = 3) +
    geom_hline(yintercept = 0.05, color = "red", linetype = 2) +
    theme_classic() +
    ylab("p-value") +
    theme(text = element_text(size = 12, family = "mono"))
```

## Permutation --> Model: step 2
```{r}
pval2 <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    fj <- suppressMessages(cmemat %>% mutate(inq = cmemat[[cmes[i]]]) %>% 
        group_by(species, aid) %>% summarize(inq = sum(inq) > 0) %>% 
        summarize(f = mean(inq)))
    pi <- left_join(hostdat, fj, by = "species") %>%
        group_by(aid) %>%
        select(aid, species, f) %>% 
        distinct %>% 
        summarize(pi = 1 - prod(1 - f)) %>%
        pull(pi)
    obs <- cmemat %>% mutate(inq = cmemat[[cmes[i]]]) %>% 
        group_by(aid) %>% summarize(present = sum(inq) > 0) %>% 
        pull(present) %>% sum
    pval2[i] <- poisbinom::ppoisbinom(obs, pi)
}
data.frame(cme = 1:length(cmes), pval = sort(pval2, decreasing = T)) %>% 
    ggplot(aes(x = cme, y = pval)) +
    geom_point(alpha = 0.8, shape = 3) +
    geom_hline(yintercept = 0.05, color = "red", linetype = 2) +
    theme_classic() +
    ylab("p-value") +
    theme(text = element_text(size = 12, family = "mono"))
```

## Permutation --> Model: step 3
```{r}
pval3 <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    print(i)
    fj <- NI[NI$cme == cmes[i], ] %>% filter(N > 0) %>% 
        mutate(f = I / N) %>% pull(f) %>% mean
    pi <- 1 - (1 - fj) ^ NI[NI$cme == cmes[i], "N"]
    obs <- cmemat %>% mutate(inq = cmemat[[cmes[i]]]) %>% 
        group_by(aid) %>% summarize(present = sum(inq) > 0) %>% 
        pull(present) %>% sum
    pval3[i] <- poisbinom::ppoisbinom(obs, pi)
}
data.frame(pval = pval3, alpha = alphas$alpha) %>% 
    arrange(-pval) %>% 
    mutate(cme = row_number()) %>% 
    ggplot(aes(x = cme, y = pval)) +
    geom_point(alpha = 0.8, shape = 3) +
    geom_hline(yintercept = 0.05, color = "red", linetype = 2) +
    theme_classic() +
    ylab("p-value") +
    theme(text = element_text(size = 12, family = "mono"))
```

## All steps
```{r}
pval0 <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    pval0[i] <- cmetab[cmetab$cseg == cmes[i], ]$p_subjects
}
pval4 <- rep(0, length(cmes))
for (i in 1:length(cmes)) {
    pval4[i] <- fits[fits$cme == cmes[i], ]$pval
}
pval_labs <- c("pval0" = "Permutation test",
               "pval1" = "Parametric version",
               "pval2" = "Aggregated at species level",
               "pval3" = "Species made exchangeable",
               "pval4" = "Model-based LRT")
data.frame(cme = cmes, 
           pval0 = pval0, pval1 = pval1, 
           pval2 = pval2, pval3 = pval3,
           pval4 = pval4,
           alpha = alphas$alpha) %>% 
    left_join(fits %>% select(cme, fu), by = "cme") %>% 
    arrange(-pval4) %>% mutate(cme = row_number()) %>% 
    pivot_longer(pval0:pval4) %>% 
    ggplot(aes(x = cme, y = value, color = fu == 0)) +
    geom_point(alpha = 0.8, shape = 3) +
    scale_color_manual(values = c("black", "grey"), 
                       labels = c("f = 0","f > 0"),
                       name = "") +
    geom_hline(yintercept = 0.05, color = "red", linetype = 2) +
    theme_classic() +
    ylab("p-value") +
    theme(text = element_text(size = 12, family = "mono")) +
    facet_wrap(~name, ncol = 1, labeller = labeller(name = pval_labs))
```

## Failure of Poisson(alpha) for data
```{r}
examineerror <- NI %>% group_by(cme) %>% 
    summarize(Ealpha = mean(N), Valpha = var(N)) %>% 
    mutate(ratio = Ealpha/Valpha)
examineerror %>%
    ggplot(aes(x = Ealpha, y = Valpha)) +
    geom_point(alpha = 0.5) +
    geom_abline(slope = 1, intercept = 0, color = "red", linetype = 2) +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))
examineerror %>% filter(ratio > 0.75) %>% select(cme, Ealpha)
```

## Significance retrodiction under fitted parameters
```{r}
sigs <- data.frame(cme = cmes, pval3 = 1, pval4 = 0.5)
idx <- 1
for (i in 1:45) {
    tmpalpha <- alphalist[i]
    counts <- suppressMessages(read_csv(paste0("alpharesults/alpha_", i, ".csv")))
    counts[, 1:O] <- counts[, 1:O] + 1 / 10
    divider <- rowSums(counts[, 1:O])
    counts[, 1:O] <- apply(counts[, 1:O], 2, \(x) x / divider)
    
    if (i == 45) { tmpcmes <- alphas$cme[alphas$alpha >= tmpalpha]
    } else tmpcmes <- alphas$cme[alphas$alpha == tmpalpha]
    
    for (j in 1:length(tmpcmes)) {
        print(idx)
        tmpN <- NI %>% filter(cme == tmpcmes[j]) %>% filter(N > 0) %>% pull(N)

        sigs[idx, 1] <- as.character(tmpcmes[j])
        
        tmppar <- fits %>% filter(cme == tmpcmes[j]) %>% 
            mutate(f = round((fl + fu) / 2, 2),
                   R0 = round((R0l + R0u) / 2, 1)) %>% 
            select(f, R0)
        
        sim <- apply((counts %>% filter(round(f, 2) == tmppar$f, round(R0, 1) == tmppar$R0))[tmpN, 1:O],
                 1, \(x)  which.max(rmultinom(1, 1, x)) - 1)
    
        statistics <- data.frame(N = tmpN, I = sim)
        
        sigs[idx, 3] <- fitting(statistics, counts)[7]
        
        fj <- statistics %>% mutate(f = I / N) %>% pull(f) %>% mean
        pi <- 1 - (1 - fj) ^ statistics$N
        if (sum(pi) > 0) {
            obs <- sum(statistics$I > 0)
            sigs[idx, 2] <- poisbinom::ppoisbinom(obs, pi)
        }
        idx <- idx + 1
    }
}
examine <- fits %>% mutate(f = round((fl + fu) / 2, 2),
                R0 = round((R0l + R0u) / 2, 1)) %>% 
    select(cme, f, R0, pval4 = pval) %>% 
    left_join(data.frame(cme = cmes, pval3 = pval3), by = "cme") %>% 
    left_join(sigs %>% rename(retro3 = pval3, retro4 = pval4), by = "cme")
examine %>%
    mutate(is0 = ifelse(R0 == 0, "R0 = 0", "R0 > 0")) %>% 
    filter(f > 0) %>% 
    ggplot(aes(x = retro4)) +
    geom_histogram(color = "black", fill = "white") + 
    facet_wrap(~is0) +
    xlab("p-value") +
    theme_classic() +
    theme(text = element_text(size = 12, family = "mono"))
```

## Sim I vs. real I
```{r}
Isim <- list()
idx <- 1
for (i in 1:45) {
    tmpalpha <- alphalist[i]
    counts <- suppressMessages(read_csv(paste0("alpharesults/alpha_", i, ".csv")))
    counts[, 1:O] <- counts[, 1:O] + 1 / 10
    divider <- rowSums(counts[, 1:O])
    counts[, 1:O] <- apply(counts[, 1:O], 2, \(x) x / divider)
    
    if (i == 45) { tmpcmes <- alphas$cme[alphas$alpha >= tmpalpha]
    } else tmpcmes <- alphas$cme[alphas$alpha == tmpalpha]
    
    for (j in 1:length(tmpcmes)) {
        print(idx)
        tmpN <- NI %>% filter(cme == tmpcmes[j]) %>% filter(N > 0) %>% pull(N)

        # sigs[idx, 1] <- as.character(tmpcmes[j])
        
        tmppar <- fits %>% filter(cme == tmpcmes[j]) %>% 
            mutate(f = round((fl + fu) / 2, 2),
                   R0 = round((R0l + R0u) / 2, 1)) %>% 
            select(f, R0)
        
        sim <- apply((counts %>% filter(round(f, 2) == tmppar$f, round(R0, 1) == tmppar$R0))[tmpN, 1:O],
                 1, \(x)  apply(rmultinom(100, 1, x), 2, which.max) - 1)

        Isim[[as.character(tmpcmes[j])]] <- sim
        
        idx <- idx + 1
    }
}
summar <- map2(Isim, names(Isim), \(x, y) {
    tmpN <- NI %>% filter(cme == y) %>% pull(N)
    tmpN <- tmpN[tmpN > 0]
    tmpstat <- apply(x, 1, \(row) mean(row / tmpN))
    c(mean(tmpstat), quantile(tmpstat, c(0.055, 0.5, 0.945)))}) %>% 
    bind_rows %>% t %>% data.frame
colnames(summar) <- c("mean","q1","q2","q3")
summar$cme <- rownames(summar)
summar <- NI %>% group_by(cme) %>% 
    summarize(f = mean(I / N, na.rm = T)) %>% 
    left_join(summar, by = "cme")
summar %>% ggplot(aes(x = f, y = q2)) +
    geom_point(alpha = 0.25) +
    geom_linerange(aes(ymin = q1, ymax = q3), alpha = 0.25) +
    geom_abline(color = "red", linetype = 2) +
    theme_classic()
problems <- summar %>% filter(q1 == 0, q2 == 0, q3 == 0) %>% pull(cme)
mean((fits %>% filter(fu == 0) %>% pull(cme)) %in% problems)
```

Essentially, when we estimate f = 0, we expect a different result

If we remove all of those CMEs, then the simulated parametric test begins to agree with the real one. However it therefore no longer agrees with the model-based test

So this explains what "data" is being lost when we simulate - which makes perfect sense
So i need to either address it or at least acknowledge it

Ok wait so this means the simulation captures the same underlying information as the data, when it comes to the parametric singificance test

which means that i can trust the retrodictive simulations and what they suggest

so now i have a decision to make

because it appeared that the parametric test was fucking the pooch in the random sample simulations

but in this retrodiction using the actual data (and not simulating from an alpha) it appears to be doing rather well, better even than the likelihood test
it's not as deep a sample but it is literally the simulations that matter - if these were the param values, as claimed, and using this conditional data, then this is what we'd find
and it's working well

so i guess the questions remains of how to deal with the f
but presumably the f is just a crunching and a mis-estimation
so we can either use a finer range of f, force it to be non-zero, or allow it to be 
and anyhow the f doesn't really matter at all bc 1. we care about R0 2. it seems we can use the other test

buttt that's still not ideal
bc our R0 plot is a lot cleaner if significance is called with the other test
in fact
it feels very reasonable
we remove most of the ones with no information
and theres a significance call it seems for anything above a certain R0 magnitude
the discrepancy is therefore just a matter of fine detail
they are fairly equivalent but not entirely equivalent in terms of how significance is actually called
visually it feels rather reasonable

